---
title: "Functions"
date: "2016-04-27T00:00:00Z"
draft: false
tags:
- R
- Statistics
summary: Statistic series
---
<script type="text/javascript" src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.1/MathJax.js?config=TeX-AMS-MML_HTMLorMML">
</script>

```{r,echo=FALSE}
options(digits=12)
```


# Exercise 1

Implement Cohen's $d$ as a function of

$$
d = f(m_1, s_1, m_2, s_2) = \frac{|m_1-m_2|}{s_{pooled}}
$$
where $s_{pooled}$ is a pooled standard deviation. Use the formula $s_{pooled} = \sqrt{(s_1^2 + s_2^2)/2}$. You may implement pooled standard deviation as a function as well.

Calculate the effect size $d$ for the differences among calories per serving, 1936 versus 2006, 1936 vs 1997 and 1997 vs 2006. Use the values from Wansink, Table 1 as given in Homework 1 or in the course outline. Name this function `cohen.d` (or similar if using SAS)

## Answer

Define your function(s) in the code chunk below, then call the function with appropriate arguments in the following sections

```{r}
# function definition
# Variables m1 and m2 are means, and s1 and s2 are standard deviations
# for two dates of comparison among calories per serving we are interested in, respectively. 
cohen.d <- function(m1,s1,m2,s2){
  cohens_d <-(abs(m1-m2)/sqrt((s1^2+s2^2)/2))
  return(cohens_d)
  }
```

#### 1936 versus 2006

```{r}
m1=268.1
m2=384.4
s1=124.8
s2=168.3
cohen.d(m1=m1,s1=s1,m2=m2,s2=s2)
```

#### 1936 versus 1997

```{r}
m1=268.1
m2=288.6
s1=124.8
s2=122.0
cohen.d(m1=m1,s1=s1,m2=m2,s2=s2)
```

#### 1997 versus 2006

```{r}
m1=288.6
m2=384.4
s1=122.0
s2=168.3
cohen.d(m1=m1,s1=s1,m2=m2,s2=s2)
```

Check your work by comparing with the previous homework.
-Answers match with previous homework!

# Exercise 2.

Implement the required replicates calculation as a function of $m_1$, $s_1$, $m_2$ and $s_2$ as required parameters, and $\alpha$ and $\beta$ as optional parameters. Let `alpha=0.05` and `beta=0.2`, so you'll need to compute quantiles for `1-alpha/2` and `1-beta`.

Your function should return an integer $n$, such that

$$
n \ge 2\times \left( \frac{CV}{\%Diff} \right)^2 \times \left(z_{\alpha/2}+ z_\beta \right)^2
$$
where $\%Diff = \frac{m_1 - m_2}{(m_1 + m_2)/2}$ and $CV = \frac{sd_{pooled}}{(m_1 + m_2)/2}$.

You may use the pooled standarad deviation function from Ex. 1 (if you defined such a function).

Name this function `required.replicates` (or similar if using SAS)

## Answer

Define your function(s) in the code chunk below, then call the function with appropriate arguments in the following sections

```{r}
# function definition
# Variables m1 and m2 are means, and s1 and s2 are standard deviations
# for two dates of comparison among calories per serving we are interested in, respectively. 
required.replicates <- function (m1,m2, s1,s2, alpha=0.05, beta=0.2){
  n <- 2* ((((sqrt((s1^2 + s2^2)/2))/(m1-m2))^2) * (qnorm((1-alpha/2)) + qnorm((1-beta)))^2) 
  return(round(n,0))
}
```
#### 1936 versus 2006

```{r}
m1=268.1
m2=384.4
s1=124.8
s2=168.3
required.replicates(m1=m1, m2=m2, s1=s1, s2=s2)
```

#### 1936 versus 1997

```{r}
m1=268.1
m2=288.6
s1=124.8
s2=122.0
required.replicates(m1=m1, m2=m2, s1=s1, s2=s2)
```

#### 1997 versus 2006

```{r}
m1=288.6
m2=384.4
s1=122.0
s2=168.3
required.replicates(m1=m1, m2=m2, s1=s1, s2=s2)
```

Check your work by comparing with the previous homework.
-Answers match with previous homework!

Note:
for Alpha=0.05 , we can use the r function qnorm(1-alpha/2) assuming u=0 and sd=1,
As for Beta, we need additional information.
z-score is the a standardized value of the value the hypothesized x.
and alpha is about rejecting the value x when  its true. 
but beta is about x failing to reject in when it is false... which means there is other value of x which we don't have in the formula z=(x-u)/sd.

# Exercise 3

Implement the likelihood formula as a function or macro. 

$$
L (x ; \mu, \sigma^2) = \frac{1}{\sigma \sqrt{2 \pi}^{}} e^{- \frac{(x - \mu)^2}{2 \sigma^2}}
$$

Define $\mu$ and $\sigma$ as optional parameters, taking values `mu=0` and `sigma=1`. Name this function `norm.pdf`
  
## Answer

Define your function(s) in the code chunk below, then call the function with appropriate arguments in the following sections

```{r}
# function definition
# Function to calcuate the values for log liklihood from above equation.
# First, we define the values for sigma as variance, 
# mu as mean of a normal population to be used for a liklihood of a x observation.
norm.pdf <- function(x,mu=0,sigma=1){
  l<-1/(sigma*sqrt(pi*2))*exp(-((x-mu)^2)/(2*sigma^2))
  return(l)
}
```

#### $x=-0.1$

```{r}
x=-0.1
norm.pdf(x)
```

#### $x=0.0$

```{r}
x=0.0
norm.pdf(x)
```

#### $x=0.1$

```{r}
x=0.1
norm.pdf(x)
```

Check your work by comparing with the previous homework.
-Answers match with previous homework!

# Exercise 4

The probability mass function for value $y$ from Poisson data with a mean and variance $\lambda$ is given by 

$$
f(x;\lambda) = \frac{e^{-\lambda} \lambda^x}{x!} = exp(-\lambda)(\frac{1}{x!}) exp[x\times log(\lambda)]
$$
Write a function `pois.pmf` that accepts two parameters, `x` and `lambda`. Use the built in `factorial` function for $x!$. Note that $x$ should be an integer value, so call a rounding function inside your function.
Test your function with $\lambda = 12$ at $x = 8,12,16$

## Answer

Define your function(s) in the code chunk below, then call the function with appropriate arguments in the following sections

```{r}
# function definition
# The function to calculate probability mass function for poisson 
# data with a mean and variance lambda. 
  pois.pmf <- function(x, lambda){
    poisson.d <- exp(-lambda)*(1/(factorial(round(x,0))))*exp(round(x,0)*(log(lambda)))
  return(poisson.d)
  }
```

#### $x=4$

```{r}
lambda=12
x=4
pois.pmf(x=x, lambda = lambda)
```

#### $x=12$

```{r}
lambda=12
x=12
pois.pmf(x=x, lambda = lambda)
```

#### $x=20$

```{r}
lambda=12
x=20
pois.pmf(x=x, lambda = lambda)
```

You can check your work against the built in Poisson distribution functions.

```{r}
# Using built-in function 'dpois', we can check our answers:
# for x=4
x= 4
lambda=12
dpois(x=x, lambda = lambda)
# #or
# ppois(x,lambda)-ppois(x-1,lambda)

# for x =12
x= 12
lambda=12
dpois(x=x, lambda = lambda)

# for x=20
x= 20
lambda=12
dpois(x=x, lambda = lambda)

# which was correct for all three x's
```

Something to ponder. Note that there are two formula given. Can you implement both forms in R/IML/Macro language? Would there be a difference in computational speed or efficiency?

```{r}
# Implementation of the first formula
pois.pmf.first <- function (x, lambda){
  poisson.d <- (exp(-lambda)*((lambda^(round(x,0))))/(factorial(round(x,0))))
  return(poisson.d)
}
# To test the execution time of two formulas:
library(microbenchmark)
lambda =12
x=20

mbm <- microbenchmark("Using first formula" = pois.pmf.first(x=x, lambda = lambda), 
                      "Using second formula" = pois.pmf(x=x, lambda = lambda))
mbm

library(ggplot2)
autoplot(mbm)
```  
Based on the execution time, looks like using the first formula takes a bit longer time to execute.

# Exercise 5

Write a function, `stat.power` that combines calculations from Exercises 1 and 2. This function should accept $m_1, s_1, m_2$ and $s_2$ as required parameters, and $\alpha$ and $\beta$ as optional parameters. This function must return a list with named elements `CV`, `PercentDiff`, `EffectSize` and `RequiredReplicates`.

If you choose to do this exercise in SAS, you will need to write a subroutine that accepts the same parameters as the R function, but also accepts output parameters `CV`, `PercentDiff`, `EffectSize` and `RequiredReplicates`. See https://blogs.sas.com/content/iml/2012/08/20/how-to-return-multiple-values-from-a-sasiml-function.html. 

Another option for SAS is to package the calculations in a macro and create a data table, using the code from 
Course Outline SAS Source (under Course Outline > Outline Source and Output Files), about line 320.


## Answer

Define your function(s) in the code chunk below, the call the function with appropriate parameters in the following sections

```{r}
# function definition
combined <- function (m1,m2, s1,s2, alpha=0.05, beta=0.2){
  n <- 2* ((((sqrt((s1^2 + s2^2)/2))/(m1-m2))^2) * (qnorm((1-alpha/2)) + qnorm((1-beta)))^2)
  cohens_d <-(abs(m1-m2)/sqrt((s1^2+s2^2)/2))
  cv <- (sqrt((s1^2+s2^2)/2))/((m1+m2)/2)
  percentdiff <- ((m1-m2)/((m1+m2)/2))
  tt <- (list(CV=cv, PercentDiff= percentdiff, RequiredReplicates=round(n,0), EffectSize=cohens_d))
  # attributes(tt)
  # names(tt)
  attr(tt, "class") <- "stat.power" #Setting a new class
  print.stat.power(tt) #use print.stat.power function below
}
```

If you define the `class` of the list returned by your function as `stat.power`, this function should work automatically; you shouln't need to call the function explicity.

```{r}
print.stat.power <- function(value) {
  cat(paste("Coefficient of Variation :",value$CV*100,"\n"))
  cat(paste("Percent Difference :",value$PercentDiff*100,"\n"))
  cat(paste("Effect Size :",value$EffectSize,"\n"))
  cat(paste("Required Replicates :",value$RequiredReplicates,"\n"))
}
```


#### 1936 versus 2006

```{r}
m1=268.1
m2=384.4
s1=124.8
s2=168.3
combined(m1=m1, m2=m2, s1=s1, s2=s2)
```

#### 1936 versus 1997

```{r}
m1=268.1
m2=288.6
s1=124.8
s2=122.0
combined(m1=m1, m2=m2, s1=s1, s2=s2)
```

#### 1997 versus 2006

```{r}
m1=288.6
m2=384.4
s1=122.0
s2=168.3
combined(m1=m1, m2=m2, s1=s1, s2=s2)
```










