---
Authors: ["**Achal Neupane**"]
title: "Data tables"
date: 2019-08-15T17:26:23-05:00
draft: false
output: html_document
tags:
- R
- Statistics
summary: Statistics series
---

<script type="text/javascript" src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.1/MathJax.js?config=TeX-AMS-MML_HTMLorMML">
</script>

# Exercise 1.

This exercise will repeat Exercise 1 from Homework 4, but using a data table.

### Part a.

Create a data table or frame with 3 defined columns:
- Let `M1` be a sequence of means from 320-420, incremented by 10. 
- Let `M2` be 270.
- Let `SD` be a pooled standard deviation of 150. 

Define and print the tabke in the space below. Do not create individual vectors for this exercise, outside of the data frame, if you use R. In SAS, you may use IML to create a matrix and save the matrix as a data table, or define a sequence (`DO`) in the DATA step. I've included framework code in the SAS template.

```{r}
M.table <- data.frame(M1= seq(320, 420, 10), M2= 270, SD = 150)
M.table
```

Add to the data table a column containing required replicates, letting $s_i = s_j = s_{pooled}$. Also add a column containing Cohen's $d$. 

To show results, either print the table or plot required replicates versus $d$ as in the previous homework.

```{r}
# Using combined function from last week to calculate effect size and Required replicates
combined <- function (m1,m2 = 270, s_pooled = 150, alpha = 0.05, beta = 0.2){
  cv <- (s_pooled)/((m1+m2)/2)
  percent.diff <- ((m1-m2)/((m1+m2)/2))
  cohens_d <-(abs(m1-m2)/(s_pooled))
  n <- 2*(((cv/percent.diff)^2)*(qnorm((1-alpha/2)) + qnorm((1-beta)))^2) 
  n <- round(n,0)
  value <- (list(CV = cv, PercentDiff= percent.diff, RequiredReplicates = round(n,0), EffectSize = cohens_d))
  return(value)
}

data <- combined(m1 = M.table$M1, s_pooled = M.table$SD)
M.table$RequiredReplicates <- data$RequiredReplicates
M.table$EffectSize <- data$EffectSize
# Wanted table
M.table

# Now plotting
# attach(M.table)
# cal.lm <- lm(M.table$RequiredReplicates ~ M.table$EffectSize)
# plot(RequiredReplicates ~ EffectSize)
plot(M.table$RequiredReplicates ~ M.table$EffectSize)
# abline(cal.lm)
abline(v = 0.5, col= 'red')
```


# Exercise 2

### Part a.
You will repeat the calculations from Homework 4, Ex 2, but this time, using a data table. However, instead of a $5 \times 6$ matrix, the result with be a table with 30 rows, each corresponding to a unique combination of CV from $8, 12, ..., 28$ and Diff from $5,10, ... , 25$.

The table should look something like

$$
\left\{
 \begin{array}{cc}
     CV & Diff \\
     8 & 5  \\
     8 & 10  \\
     8 & 15  \\
     \vdots & \vdots \\
     12 & 5  \\
     12 & 10  \\
     12 & 15  \\
     \vdots & \vdots \\
     28 & 5  \\
     28 & 10  \\
     28 & 15  \\
   \end{array}
   \right\}
$$
Test your required replicates calculations by calculating required replicates for each combination of CV and Diff using the default values for $\alpha$ and $\beta$. Name this column `Moderate`. 

Calculate required replicaes again, but this time let $\alpha = 0.01$ and let $\beta = 0.1$. Label this column `Conservative`. Repeat the calculations, but this time let $\alpha = 0.1$ and let $\beta = 0.2$.

If you choose SAS, you can use the framework code from the first exercise.

```{r}
newdata.dat <- data.frame(CV = rep(seq(8,28,4), each = 5), Diff = rep(seq(5,25,5),6))
newdata.dat

combined <- function (cv, percent.diff, alpha = 0.05, beta = 0.2){
  cv <- cv
  percent.diff <- percent.diff
  n <- 2*(((cv/percent.diff)^2)*(qnorm((1-alpha/2)) + qnorm((1-beta)))^2) 
  n <- round(n,0)
  value <- list(CV = cv, PercentDiff= percent.diff, RequiredReplicates = round(n,0))
  return(value)
}
value <- combined(cv = newdata.dat$CV, percent.diff = newdata.dat$Diff)

# Adding Moderate column
newdata.dat$Moderate <- value$RequiredReplicates

# Adding Conservative column
value <- combined(cv = newdata.dat$CV, percent.diff = newdata.dat$Diff, alpha = 0.01, beta = 0.1)

newdata.dat$Conservative <- value$RequiredReplicates

# Repeat the calculations, but this time let $\alpha = 0.1$ and let $\beta =
# 0.2$.

# Adding Liberal column
value <- combined(cv = newdata.dat$CV, percent.diff = newdata.dat$Diff, alpha = 0.1, beta = 0.2)

newdata.dat$Liberal <- value$RequiredReplicates
# Print the table
# newdata.dat
```

To show your work, some ideas for graphs:

- Plot required replicates vs CV and Diff, using different colors or symbols for `Moderate`,`Conservative` and `Liberal` 
```{r}
attach(newdata.dat)

# Effect size Vs Required replicates
plot(CV, Moderate, col = "black", type = 'b', pch = 8, xlab="Effect Size", ylab="Required replicates", ylim = c(min(Moderate, Conservative, Liberal),max(Moderate, Conservative, Liberal)))
points(CV, Conservative, type = 'b', pch = 2,  col = "red")
points(CV, Liberal, type = 'b', pch = 20,  col = "blue")
legend("topleft", 
       legend = c("Moderate", "Conservative", "Liberal"), 
       col = c("black", "red", "blue"),
       bty = "o", 
       pch = c(8, 2, 20),
       pt.cex = 2, 
       cex = 1.2, 
       text.col = "black", 
       horiz = F , 
       inset = c(0.1, 0.1))

# Percent diff vs Required replicates
plot(Diff, Moderate, col = "black", type = 'b',pch = 8, xlab="Percent diff", ylab="Required replicates", ylim = c(min(Moderate, Conservative, Liberal),max(Moderate, Conservative, Liberal)))
points(Diff, Conservative, type = 'b', pch = 2,  col = "red")
points(Diff, Liberal, type = 'b', pch = 20,  col = "blue")
legend("topright", 
       legend = c("Moderate", "Conservative", "Liberal"), 
       col = c("black", "red", "blue"),
       bty = "o", 
       pch = c(8, 2, 20),
       pt.cex = 2, 
       cex = 1.2, 
       text.col = "black", 
       horiz = F , 
       inset = c(0.1, 0.1))
```

- Plot `Conservative` vs `Moderate` and `Liberal` vs `Moderate`, including a line with slope 1 and intercept 0.

```{r}
attach(newdata.dat)
#Conservative Vs Moderate
plot(Conservative ~ Moderate, type = 'b', col = "black")
abline(a=1, b = 0)
#Liberal Vs Moderate
plot(Liberal ~ Moderate, type = 'b', col = "black")
abline(a=1, b = 0)
```


# Exercise 3

You'll work with data from U.S. Wholesale price for pumpkins 2018 (https://www.ers.usda.gov/newsroom/trending-topics/pumpkins-background-statistics/, Table 1)

## Part a

Download the file `pumpkins.csv` from D2L and read the file into a data frame. Print a summary of the table.

```{r}
pumpkins <- read.csv("/Users/owner1/Box/sdsu/statistical_programming_course/Week5/pumpkins.csv", header = TRUE, sep = ",")
pumpkins <- data.frame(pumpkins)
attach(pumpkins)
print(summary(pumpkins))
```

## Part b
To show that the data was read correctly, create three plots. Plot
1. Price vs Week 
2. Price vs Class
3. Size vs Class

These three plots should reproduce the three types of plots shown in the `RegressionEtcPlots` video, **Categorical vs Categorical**, **Continuous vs Continuous** and **Continuous vs Categorical**. Add these as titles to your plots, as appropriate.

```{r}
attach(pumpkins)
plot(Price ~ Week, main = "**Continuous vs Continuous**")
plot(Price ~ Class, main = "**Continuous vs Categorical**")
plot(Size ~ Class, main = "**Categorical vs Categorical**")
```

From these plots, you should be able to answer these questions:

1. Are some Weeks missing Price observations?
Yes some Weeks are missing some Price observations.

2. Do Prices vary more for some Classes?
Yes, prices vary more for some classes, such as Pie.

2. Do all Classes have the same Sizes?
No, they are of different sizes.

# Exercise 4

Calculate a one-way analysis of variance from the pumpkin data in Exercise 3. 

## Option A

Let $y$ be the `Price`. Let the $k$ treatments be `Class`. Let $T_i$ be the `Price` total for `Class` $i$ and let $r_i$ be the number of observations for `Class` $i$. Denote the total number of observations $N = \sum r_i$.

### Part a
Find the treatment totals $\mathbf{T} = T_1 \dots T_k$ and replicates per treatment $\mathbf{r} = r_1 \dots r_k$ from the pumpkin data, using group summary functions and compute a grand total $G$ for `Price`. Print $\mathbf{T}$, $\mathbf{r}$ and $G$ below. In SAS, you can use `proc summary` or `proc means` to compute $T$ adn $r$ and output a summary table. I find the rest is easier in IML (see `use` to access data tables in IML).

```{r}
pumpkins.dat <- read.table("/Users/owner1/Box/sdsu/statistical_programming_course/Week5/pumpkins.csv", header = TRUE, sep = ",")
head(pumpkins.dat)
class(pumpkins.dat)
k <- length(levels(pumpkins.dat$Class))
k

T <- tapply(pumpkins.dat$Price, pumpkins.dat$Class, sum)
# can also use aggregate 
# T <- aggregate(pumpkins.dat$Price, by = list(pumpkins.dat$Class), FUN = sum)
T
r <- table(pumpkins.dat$Class)
# can also use aggregate
# aggregate(pumpkins.dat$Class, by = list(pumpkins.dat$Class), FUN = length)
r
N <- sum(r)
G <- sum(pumpkins.dat$Price)
G
```

### Part b
Calculate sums of squares as

$$
\begin{aligned}
\text{Correction Factor : } C &= \frac{G^2}{N} \\
\text{Total SS : } &= \sum y^2 - C \\
\text{Treatments SS : }  &= \sum \frac{T_i^2}{r_i} -C \\
\text{Error SS : }  &= \text{Total SS} - \text{Treatments SS} \\
\end{aligned}
$$
and calcute $MSB = (\text{Treatments SS})/(k-1)$ and $MSW = (\text{Error SS})/(N-k)$.

```{r}
C <- G^2/N
C
TotalSS <- sum((pumpkins.dat$Price)^2)-C
TotalSS
TreatmentsSS <- sum(T^2/r) - C
TreatmentsSS
ErrorSS <- TotalSS - TreatmentsSS
ErrorSS
MSB <- TreatmentsSS/(k-1)
MSB
MSW <- ErrorSS/(N-k)
MSW
```

### Part c.
Calculate an F-ratio and a $p$ for this $F$, using the $F$ distribution with $k-1$ and $N-k$ degrees of freedom. Use $\alpha=0.05$. 

```{r}
F.ratio <- MSB/MSW
F.ratio
df1 <- k-1
df2 <- N-k
p.value <- pf(F.ratio, df1, df2, lower.tail = FALSE)
p.value
```

To check your work, use `aov` as illustated in the chunk below:

```{r,eval=TRUE}
summary(aov(Price ~ Class, data=pumpkins.dat))
```

## Option B
You may resue code from Exercise 6, Homework 4. Use group summary functions to calculate means, standard deviations and replicates from the pumpkin data, then calculate $MSW$ and $MSB$ as previously. Report the F-ratio and $p$ value as above.

```{r}
mean.groups <- aggregate(pumpkins.dat$Price, by = list(pumpkins.dat$Class), FUN = mean)
mean.groups
mean.groups <- mean.groups$x
sd.groups <- aggregate(pumpkins.dat$Price, by = list(pumpkins.dat$Class), FUN = sd)
sd.groups
sd.groups <- sd.groups$x

k <- length(mean.groups)
k

length.groups <- aggregate(pumpkins.dat$Price, by = list(pumpkins.dat$Class), FUN = length)

# We have n samples in each pumpkin.dat Class; so our population size is N:
n <- length.groups$x
n
N <- sum(n)
N
#calculating MSB 
mean.mean.groups <- mean(mean.groups)
MSB = (sum(n*(mean.groups-mean.mean.groups)^2))/(k-1)
MSB

MSW <- sum((n-1) * sd.groups^2)/(N-k)
MSW

F.ratio <- MSB/MSW
F.ratio
df1 <- k-1
df2 <- N-k
p.value <- pf(F.ratio, df1, df2, lower.tail = FALSE)
p.value
```

# Exercise 5

## Part a
Go to http://www.itl.nist.gov/div898/strd/anova/SiRstv.html and use the data listed under `Data File in Table Format` (https://www.itl.nist.gov/div898/strd/anova/SiRstvt.dat)


## Part b
Edit this into a file that can be read into R or SAS, or find an appropriate function that can read the file as-is. You will need to upload the edited file to D2L along with your Rmd/SAS files. Provide a brief comment on changes you make, or assumptions about the file needed for you file to be read into R/SAS. Read the data into a data frame or data table.

```{r}
# Here we read the .dat file first skipping the 59 lines as indicated in SiRstvt.dat file.
df <- read.table("/Users/owner1/Box/sdsu/statistical_programming_course/Week5/SiRstvt.dat", header = FALSE, skip = 59)
# read file as dataframe
df <- as.data.frame(df)
# We need to change the column names as alpha-numeric to work with ease
names(df) <- paste0("col_", seq(1,5,1))
df
```

## Part c

There are 5 columns in these data. Calculate mean and sd and sample size for each column in this data, using column summary functions. Print the results below

```{r}
# Calculating the mean (or summary)
sapply(df, mean)
# summary(df, digits = 7)
print(data.frame(sapply(df, summary)))

# calculating length for all columns
sapply(df, length)

# calculating sd for all columns
sapply(df, sd)
```

Determine the largest and smallest means, and their corresponding standard deviations, and calculate an effect size and required replicates to experimentally detect this effect.

If you defined functions in the previous exercises, you should be able to call them here.

```{r}
tt<- as.data.frame(sapply(df, summary))
# Assigning smallest mean as m1 (Col_5)
m1 <- min(tt[rownames(tt)=="Mean",])
m1
# Assigning largest mean as m2 (col_2)
m2 <- max(tt[rownames(tt)=="Mean",])
m2
# Now, getting sd for column 2 (largest mean) and 5 (smallest mean)
sd <- sapply(df, sd)
# SD from smallest mean column and largest mean column
# SD for column 5 (smallest mean)
s1 <- as.numeric(sd[5])
s1
# SD for column 2 (largest mean)
s2 <- as.numeric(sd[2])
s2

# Calculating the effect size
cohen.d <- function(m1,s1,m2,s2){
  cohens_d <-(abs(m1-m2)/sqrt((s1^2+s2^2)/2))
  return(cohens_d)
}

cohen.d(m1 = m1, s1 = s1, m2 = m2, s2 = s2)

# calculating the Required replicates
required.replicates <- function (m1,s1, m2,s2, alpha=0.05, beta=0.2){
  n <- 2* ((((sqrt((s1^2 + s2^2)/2))/(m1-m2))^2) * (qnorm((1-alpha/2)) + qnorm((1-beta)))^2) 
  return(round(n,0))
}

# required replicates
required.replicates(m1 = m1, s1 = s1, m2 = m2, s2 = s2)
```

# Exercise 6

There is a web site (https://www.wrestlestat.com/rankings/starters) that ranks college wrestlers using an ELO scoring system (https://en.wikipedia.org/wiki/Elo_rating_system). I was curious how well the rankings predicted performance, so I gathered data from th 2018 NCAA Wrestling Championships (https://i.turner.ncaa.com/sites/default/files/external/gametool/brackets/wrestling_d1_2018.pdf). Part of the data are on D2L in the file `elo.csv`. You will need to download the file to your computer for this exercise.

Read the dzta below and print a summary. The dzta were created by writing a data frame from R to csv (`write.csv`), so the first column is row number and does not have a header entry (the header name is an empty string).

```{r}
elo.dat <- read.table("/Users/owner1/Box/sdsu/statistical_programming_course/Week5/elo.csv", header = TRUE, row.names = 1, sep = ",")
# elo.dat
print(summary(elo.dat))
```

Each row corresponds to an individual wrestler, his weight class and collegiate conference. The WrestleStat ELO score is listed, along with his tournament finish round (i.e. `AA` = 1-8 place, `cons 12` = lost in the final consolation round, etc.). I calculated an expected finish based on his ELO ranking within the weight class, where `E[AA]` = top 8 ranked, expected to finish as AA, etc.

Produce group summaries or plots to answer the following:

- What are the mean and standard deviations of ELO by Expected Finish and by Actual Finish?
```{r}
# By both groups
# mean.ExActFinish <- aggregate(elo.dat$ELO, by = list(elo.dat$ExpectedFinish, elo.dat$ActualFinish), mean)
# mean.ExActFinish

# First By Expected Finish:
mean.ExFinish <- aggregate(elo.dat$ELO, by = list(elo.dat$ExpectedFinish), FUN = mean)
mean.ExFinish

SD.ExFinish <- aggregate(elo.dat$ELO, by = list(elo.dat$ExpectedFinish), FUN = sd)
SD.ExFinish

# Now, by Actual Finish
mean.ActFinish <- aggregate(elo.dat$ELO, by = list(elo.dat$ActualFinish), FUN = mean)
mean.ActFinish

SD.ActFinish <- aggregate(elo.dat$ELO, by = list(elo.dat$ActualFinish), FUN = sd)
SD.ActFinish
```

- Do all conferences have similar quality, or might we suspect one or more conferences have better wrestlers than the rest? (You don't need to perform an analysis, just argue, based on the summary, if a deeper analysis is warranted).
```{r}
aggregate(elo.dat$Conference, by = list(elo.dat$ExpectedFinish, elo.dat$ActualFinish), FUN = summary)
# aggregate(elo.dat$Conference, by = list(elo.dat$ExpectedFinish, elo.dat$ActualFinish), length)
```
Based on this table, Big Ten seems to have better wrestlers.


- How well does ELO predict finish? Use a contingency table or mosaic plot to show how often, say, and `AA` finish corresponds to an `E[AA]` finish.
```{r}
contingency.table <- table(elo.dat$ExpectedFinish, elo.dat$ActualFinish)
# Contingency table
contingency.table
# # can convert table to dataframe
# contingency.table.dat <- as.data.frame.matrix(contingency.table)
# Also, plot
attach(elo.dat)
plot(ActualFinish ~ ExpectedFinish)
```
Based on the contingency table, `E[AA]` `AA` are associated 57 times 

- Does this data set include non-qualifiers? (The NCAA tournament only allows 33 wreslers per weight class).
```{r}
table(elo.dat$Weight)
```
Based on this, all weight class have 33 wrestlers with only 32 wrestlers in 197 weigh class
