---
Authors: ["**Achal Neupane**"]
title: "Final assignment"
date: 2019-09-15T17:26:23-05:00
draft: false
output: html_document
tags:
- R
- Statistics
- Statistical_programming
summary: Statistics series
---

<script type="text/javascript" src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.1/MathJax.js?config=TeX-AMS-MML_HTMLorMML">
</script>

**Project Overview:**

The final project will be a continuation of the midterm project. We will continue working with data from four corn fields, but will be looking at some issues with times and dates.
For each field, I've uploaded two files, seed*.csv and harvest*.csv, corresponding to seeding rate and yield data, respectively. Both sets of files have a column Timestamp with strings of the form
2018-05-20T13:20:08.201Z
Your task will be to extract the date and time values from these strings, and answer these questions. The text before 'T" is the date string, and the text between 'T" and "Z" is the time string, in universal time.
1. What is the range of planting dates in these data?
2. Was each field harvested entirely in one day, and where they harvested each at approximately the same time of day?
It might be enough to process only the first and last rows in the data. Each should be sampled at one second intervals, so the time difference between the first and last rows, in seconds, should be almost equal to the number of rows. I would not be too worried about gaps on the data on the order of seconds, but I would want to know about hour long gaps in the data, and where they occur. Note that time will reset to 0 at 23:59:59 and date will increment, so account for this in the analysis.
Some additional thoughts. 
There is a relationship between planting date and yield. We could review the literature to get a more precise estimate (and you can do that, if you wish), but we'll start with a rough back-of-the-envelope calculation. 
Suppose we are working with 100-day corn - we expect corn to reach maturity in roughly 100 days. Let's suppose that the difference between the first field planting date and the last field planting date is 5 days. That's 5 percent of the growing period, so let %Diff = 5. What is the standard deviation for Yield, with regard to planting date? Again, we can look to the literature, but we'll use a simplifying assumption that it will be similar to the sd for yield vs planting rate, as determined in the midterm project. Convert that to CV. 
Now we have a first approximation for effect size (%Diff/CV). Is this effect size large enough that we need to worry that the effect of planting date will confound our analysis of the relationship between yield and planting rate? When I gather data from more fields, do I need to be careful and only include fields in a narrow range of dates? What is the first approximation for the number of fields required to test the relationship between planting date and yield?
Not, about gaps in harvest. I've been arguing that analyzing yield monitor data should be a two-step process. First, analysis the yield as a time series (as grain moves through the harvester) then analyze as spatial data (the as the harvester moves over the field). There will be auto-correlated errors in each process that should be analyzed independently.
In particular, the value for yield as reported in these data is not the yield of the grain as it is measured going through the harvester. The grain moving through the harvester will be of varying degrees of maturity, thus of grain moisture - less mature grain will have more water content, thus more weight. Yield values are standardized to a define percent moisture, so the harvester has a moisture sensor, and the percent moisture reading is use to normalize yield.
But yield is measured at one second intervals, while moisture can only be measured at approximately 10-15 second intervals. I'm curious if gaps in the harvest record will affect how yield is normalized by moisture - there may be some cases where the moisture reading is uncorrelated with yield, because of this difference in measurement (I suspect it will be very small, but it's an interesting problem, to me).
I'm also interested in how percent moisture changes over the course of a day, and if there is an effect when fields are harvested at different times of day.
More will follow in the discussions.

**Solution**


The algorithm to do this analysis is divided into several steps as described below:

1. First you have to define the path to all csv files for final under (`path.final`) and all .csv files from midterm under `path.midterm` in the begining of the code below.
2. I have used some external packages like ggplot2 to make the plots.
3. I have written several functions to analyze each individual field as well as merged field statistics that calculates cohens effect size, required replicates, and does ANOVA analysis on individual field as well as on merged data from all fields. 
4. This function `get_question_1_and_2_answers` calculates the time spent each day on field, start and end time start and end date. observations each day, and also calculate one hour gap (more than an hour gap in consecutive rows) if present in the data. It spits out a table (`range_table`) with all the date data required for this project. I then used this range table to plot time, and dates.
5. I have also plotted the Moisture content in each harvest field and how it gets affected during the day, and across the range of dates.
6. Function `calculate_field_mean_SD_and_get_RR_EffSize` also uses mean, sd, and counts from each field to calculate cohens d and Required replicates. It also uses pooled standard deviations and Mean (mean of Means) calculated from all four fields for different control Rates to calculate cohens d and required replicates of combined field data. I have tested this function on all four fields at ControlRate level of 500, 1000, 2000 and 3000 intervals for mid-term, but this time, I have modified this function to calculate cohends d and required replicates for 1000 control rates also taking dates into consideration. This function also calculates these values for date only (neglecting control Rates).
7. I have also shown ANOVA analysis followed by TUKEY HSD test to show which ControlRates of seeding have significant effect on Yield for each field and then for all fields (combined all four fields).
8. Lastly, I have plotted EffectSize Vs RequiredReplicates for all combined field.

First of all we, will set two directory paths where we have all our files we want to anlalyze. first path is `path.final` for final data and `path.midterm` for midterm data

`all.files` contains the file name of all csv files in our directory. I have decided to
read all files from the system rather than reading one file at a time.

```{r}
# # Path to all .csv data files for final.
# path.final <-
# "/Users/owner1/Box/sdsu/statistical_programming_course/final/"
# 
# # Path to all .csv data files for midterm.
# path.midterm <-
#   "/Users/owner1/Box/sdsu/statistical_programming_course/midterm/"

# instead for local path, we can get files from git repos
library(rvest)
path.final <- read_html("https://github.com//achalneupane/data") %>% html_nodes(".js-navigation-open") %>% html_attr("href")

path.midterm <- read_html("https://github.com//achalneupane/data") %>% html_nodes(".js-navigation-open") %>% html_attr("href")
```

Now, we first install and load some of the packages  ("multcompView", "ggplot2",
"scales", "data.table", etc.) we will be using for this exercise.

```{r}
# First, install missing packages and load them
myPackages <-
c(
"multcompView",
"ggplot2",
"scales",
"data.table",
"reshape2",
"RColorBrewer",
"plyr",
"ggpmisc"
)
my.installed.packages <- installed.packages()
available.packages <- myPackages %in% my.installed.packages
if (sum(!available.packages) > 0) {
install.packages(myPackages[!available.packages])
}
# Load all required packages
lapply(myPackages, require, character.only = TRUE)
```

Normally, I would write all functions in a separate file and use them as
source('function_file.r') However, I have my functions for this project written
here.
```{r}
# field_or_seed <- harvestB


# functions

# function for question 1 and 2
get_question_1_and_2_answers <- function(field_or_seed) {
# range of planting dates
field_or_seed$Date <- gsub("T.*", "", field_or_seed$Timestamp)
field_or_seed$Time <- gsub(".*T|Z", "", field_or_seed$Timestamp)
# Range in days
field_or_seed$date_time <-
as.POSIXct(paste(field_or_seed$Date, field_or_seed$Time), tz = "UTC")
# field_or_seed$date_time <- paste(field_or_seed$Date, field_or_seed$Time)

# num_days <- round(difftime(max(as.POSIXct(field_or_seed$date_time)), min(as.POSIXct(field_or_seed$date_time))), 3)
num_days <-
round(difftime(max(as.POSIXct(
field_or_seed$date_time
)), min(as.POSIXct(
field_or_seed$date_time
)), units = "days"), 3)

range_from_start_to_end_date_in_days <-
ifelse(num_days <= 1,
paste0(num_days, " day"),
paste0(num_days, " days"))
# Range in span of date for the field
range_of_start_end_date <-
paste0(range(as.Date(field_or_seed$Date))[1], " to ", range(as.Date(field_or_seed$Date))[2])



field_or_seed$Date <- as.factor(field_or_seed$Date)

# time range each day
tt <- field_or_seed[, c("Date", "Time")]




time.range.each.day <- aggregate(data.frame(Time = strptime(do.call(paste, tt), '%F %R:%OS', tz = 'UTC')),
by = list(Date = tt$Date),
function(Time) {
paste(format(min(Time), '%T'),
format(max(Time), '%T'),
sep = ' to ')
})

rm(tt)


# Function to find more than an hour gap in consecutive rows
about_an_hour_gap <- function(field_or_seed) {
one_hour_thing <- {
}
for (i in 1:nrow(field_or_seed)) {
# i=3
if (i == nrow(field_or_seed))
{
break
} else
time.lapse <-
abs(as.numeric(
difftime(field_or_seed$date_time[i + 1], field_or_seed$date_time[i]),
units = "mins"
))

# find consecutive rows where the time is more than or equal 60 minutes and
# dates are the same
if (time.lapse >= 60 &
field_or_seed$Date[i + 1] == field_or_seed$Date[i]) {
one_hour_thing.tmp <- field_or_seed[i:(i + 1), ]
one_hour_thing <- rbind(one_hour_thing, one_hour_thing.tmp)
}
}
return(one_hour_thing)
}
about_an_hour_gap(field_or_seed)
one_hour_gap_time_stamp <- about_an_hour_gap(field_or_seed)

# function to calculate total time spent each day
time.elapsed.each.day.function <- function(x) {
difftime(max(as.POSIXct(x)), min(as.POSIXct(x)), units = "mins")
}


# time.range.each.day(tt)
time.elapsed.each.day <-
aggregate(date_time ~ Date, data = field_or_seed, FUN = time.elapsed.each.day.function)
colnames(time.elapsed.each.day) <-
c("Date", "Time_elapsed (minutes)")

observations.each.day <- table(field_or_seed$Date)


# Elapsed Dates
dates.elapsed.each.field.function <- function(x){
  difftime(max(as.POSIXct(x)), min(as.POSIXct(x)), units = "days")
}
elapsed.dates.each.field <- as.integer(dates.elapsed.each.field.function(field_or_seed$Date))

return(
list(
observations.each.day = observations.each.day,
range_from_start_to_end_date_in_days = range_from_start_to_end_date_in_days,
range_of_start_end_date = range_of_start_end_date,
time.elapsed.each.day = time.elapsed.each.day,
elapsed.dates.each.field = elapsed.dates.each.field, 
time.range.each.day = time.range.each.day,
one_hour_gap_time_stamp = one_hour_gap_time_stamp
)
)


}



# Functions to calculate cohen's d and Required Replicates
cohen.d <- function(m1, s1, m2, s2) {
cohens_d <- (abs(m1 - m2) / sqrt((s1 ^ 2 + s2 ^ 2) / 2))
return(cohens_d)
}

required.replicates <-
function (m1,
s1,
m2,
s2,
alpha = 0.05,
beta = 0.2) {
n <-
2 * ((((sqrt((s1 ^ 2 + s2 ^ 2) / 2
)) / (m1 - m2)) ^ 2) * (qnorm((1 - alpha / 2)) + qnorm((1 - beta))) ^ 2)
return(round(n, 0))
}


# Additional functions: They do anova and also calculate required replicates for fields
# We willl also perform ANOVA analysis with Tukey Test for paired comparision of
# mean for each field data as well as merged data at different ControlRate
# intervals. This function does Tukey HSD test and generates label for
# significant outcomes.

# Create function to get the labels for Tukey HSD:
generate_label_df <- function(TUKEY, variable) {
# Extract labels and factor levels from Tukey post-hoc
Tukey.levels <- TUKEY[[variable]][, 4]
Tukey.labels <-
data.frame(multcompLetters(Tukey.levels)['Letters'])

#I need to put the labels in the same order as in the boxplot :
Tukey.labels$treatment = rownames(Tukey.labels)
Tukey.labels = Tukey.labels[order(Tukey.labels$treatment) ,]
return(Tukey.labels)
}

# This function does ANOVA and makes boxplots with Tukey statistics for
# comparing Mean yield.
get_my_box_plot <- function (field, plot_name = "the Field") {
field$CR.Date.Levels <- gsub("-", "_", field$CR.Date.Levels)
field$CR.Date.Levels <- as.factor(field$CR.Date.Levels)
model = lm(field$Yield ~ field$CR.Date.Levels)
ANOVA = aov(model)
# If residual degrees of freedom is less than or equal 1, don't do Tukey test;
# simply return ANOVA summary.
if (ANOVA$df.residual <= 1) {
return(summary.aov(ANOVA))
}
# Tukey test to study each pair of treatment :
TUKEY <-
TukeyHSD(x = ANOVA,
'field$CR.Date.Levels',
conf.level = 0.95)

# generate labels using function
labels <- generate_label_df(TUKEY , "field$CR.Date.Levels")

# rename columns for merging
names(labels) <- c('Letters', 'CR.Date.Levels')

# Obtain letter positions for y axis using means
yvalue <- aggregate(Yield ~ CR.Date.Levels, data = field, mean)

final <- merge(labels, yvalue) #merge dataframes

p <- ggplot(field, aes(x = CR.Date.Levels, y = Yield)) +
geom_blank() +
theme_bw() +
# theme(panel.grid.major = element_blank(),
# panel.grid.minor = element_blank()) +
labs(x = 'CR.Date.Levels', y = 'Mean Yield') +
ggtitle(paste0("CR.Date.Levels Vs Mean yield for ", plot_name),
expression(atop(italic(
"(Anova:TukeyHSD)"
), ""))) +
# ggtitle(paste0("CR.Date.Levels Vs Mean yield for ", plot_name)) +
theme(plot.title = element_text(hjust = 0.5, face = 'bold')) +
geom_boxplot(fill = 'grey', stat = "boxplot") +
# coord_cartesian(clip = 'off') +
geom_text(
data = final,
aes(x = CR.Date.Levels, y = Yield, label = Letters),
vjust = -3.5,
hjust = -.5
) +
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
# geom_vline(aes(xintercept = 4.5), linetype = "dashed") +
theme(plot.title = element_text(vjust = -0.6)) +
coord_cartesian(ylim = c(min(field$Yield), max(field$Yield) + 10))
print(p)
}
```

Function to calculate mean and sd for each field for a given interval of
ControlRate and then calculate Required Replicates and EffectSize. I have made
one function to do all that so I can just use this function to analyze all
four fields. To analyze single field set `Single.Field.Analysis = TRUE`, and 
to analyze all four field together, set `Single.Field.Analysis = FALSE`,

```{r}
calculate_field_mean_SD_and_get_RR_EffSize <- 
  function(
    field, intervals = 1000, Single.Field.Analysis = TRUE, Date.Only = FALSE, printplot = FALSE){
    options(warn=1)
    plot_NAME <- (deparse(substitute(field)))
    if(Single.Field.Analysis > 0){
      # removing rows with NAs in Yield
      sum(is.na(field$Yield))
      # field <- field[!complete.cases(field$Yield), ]
      field <- field[!is.na(field$Yield),]
      # as.factor(fieldA$ControlRate)
      
      
      # work on timestamp columns
      field$Date <- gsub("T.*", "", field$Timestamp)
      field$Time <- gsub(".*T|Z", "", field$Timestamp)
      # Range in days
      field$date_time <- as.POSIXct(paste(field$Date, field$Time), tz = "UTC")
      
      
      # table.field <- table(as.factor(field$Date))
      field$ControlRate.Levels <- as.factor(intervals * ceiling(field$ControlRate/intervals))
      if(Date.Only < 1){
      field$CR.Date.Levels <- as.factor(paste(field$ControlRate.Levels, field$Date, sep = "*"))
      } else {
        field$CR.Date.Levels <- as.factor(field$Date)
      }
      table.field <- table(as.factor(field$CR.Date.Levels))
      
      
      
      
      field.Count <- setNames(aggregate(field$CR.Date.Levels, 
                                        by = list(field$CR.Date.Levels), FUN = length), c("CR.Date.Levels", "Count"))
      # # Degree of freedome = n * k - k
      # field.Count$degree.Freedom <- 
      #   (field.Count$Count * length(field.Count$ControlRate)) -length(field.Count$ControlRate)
      
      field.mean <-
        setNames(aggregate(
          field$Yield,
          by = list(field$CR.Date.Levels),
          FUN = mean
        ),
        c("CR.Date.Levels", "Mean"))
      
      
      field.SD <-
        setNames(aggregate(
          field$Yield,
          by = list(field$CR.Date.Levels),
          FUN = sd
        ),
        c("CR.Date.Levels", "SD"))
      
      
      if(printplot == 1 ){
        get_my_box_plot(field, plot_name = plot_NAME)
      }
      
      # plot individual fields with tukey test We will print box plot only if we
      # want for certain ControlRates intervals. Otherwise we will have too many
      # plots
    } else {
      temp.Field <- field
      colnames(temp.Field)[colnames(temp.Field) == "CR.Date.Levels"] <- "ControlRate.Levels"
      colnames(temp.Field)[colnames(temp.Field) == "Mean"] <- "Yield"
      # get_my_box_plot(temp.Field)
      field.SD  <- 
        as.data.frame(cbind(ControlRate = field["CR.Date.Levels"], SD = field["SD_pooled"]))
      field.mean <- 
        as.data.frame(cbind(ControlRate = field["CR.Date.Levels"], Mean = field["Mean"]))
    }
    
    # Calculate Required replicate and Effect Size from each
    # field for ControlRate i vs i+1
    
    # ReqRep_EffectSize_table <- function (field.mean, field.SD){
    Req.Rep.table.field <- {}
    for (i in 1:nrow(field.SD)){
      if(i+1 > nrow(field.SD) ){
        break
      }
      temp.Effect.size <-
        cohen.d(
          m1 = field.mean$Mean[i],
          s1 = field.SD$SD[i],
          m2 = field.mean$Mean[i + 1],
          s2 = field.SD$SD[i + 1]
        )
      
      tmp.req.reps <-
        required.replicates(
          m1 = field.mean$Mean[i],
          s1 = field.SD$SD[i],
          m2 = field.mean$Mean[i + 1],
          s2 = field.SD$SD[i + 1]
        )
      
      tmp.table <-
        cbind(
          Group = paste0(field.SD$CR.Date.Levels[i], " Vs ", field.SD$CR.Date.Levels[i + 1]),
          EffectSize = temp.Effect.size,
          RequiredReplicates = tmp.req.reps
        )
      
      Req.Rep.table.field <- rbind(Req.Rep.table.field, tmp.table)
    }
    if (Single.Field.Analysis > 0) {
      return(
        list(
          field.mean = field.mean,
          fieldSD = field.SD,
          field.Count = field.Count,
          Req.Rep.table.field = Req.Rep.table.field
        )
      )
    } else{
      return(list(Req.Rep.table.field = Req.Rep.table.field))
    }
  }
```

Here, I am reading all files in my directory we set in the begining
(path.final). I was being a bit lazy to copy hard coded paths. I then read those
files in for loop and assigned them to their respective variable names [eg,
gsub(".csv","",filename.csv)].

```{r}
# all.files contains the file name of all csv files in our directory. I have decided to
# read all files from the system rather than reading one file at a time.

# setwd(path.final)
# if using local path
# all.files <- list.files(path.final)

# using from github
all.files <- path.final
all.files <- all.files[grepl("*.csv", all.files)]

# # selecting only required files


all.files <- all.files[grepl("seedA.csv|seedB.csv|seedC.csv|seedD.csv|harvestA.csv|harvestB.csv|harvestC.csv|harvestD.csv", all.files)]

# # Or choose specifically which data files in path to analyze
# all.files <- all.files <- c("seedA.csv", "seedB.csv", "seedC.csv", "seedD.csv")

# Rearranging the vectors, putting seed data first and then harvest data
all.files <-
c(all.files[grepl("seed", all.files)], all.files[grepl("harvest", all.files)])
# all.files
# [1] "harvestA.csv" "harvestB.csv" "harvestC.csv" "harvestD.csv" "seedA.csv"    "seedB.csv"    "seedC.csv"    "seedD.csv"

#https://raw.githubusercontent.com/achalneupane/data/master/
# reading all csv files within the path and saving them to their respective names.
for (i in 1:length(all.files)) {
assign(
gsub(".*master/|.csv", "", all.files)[i],
read.table(gsub("blob/", "",paste0("https://raw.githubusercontent.com",all.files[i])),
header = TRUE,
sep = ",")
)
print(paste0("Read file ", all.files[i], ' !'))
}
# head(harvestA)
# head(harvestB) # and so on..
```

Here, I am using `get_question_1_and_2_answers` to extract date information by
looping over all data files we just read above.

```{r}
# make a table of date range
range_table <- {}
one.hour.time.difference <- data.frame()
for (i in 1:length(all.files)) {
# file.dat <- eval(sub(".csv", "", all.files, fixed = TRUE)[i])
file.dat <- eval(gsub(".*master/|.csv", "", all.files)[i])
# print(paste0("Doing time thing for ", file.dat, "..."))
dat <- get_question_1_and_2_answers(eval(parse(text = file.dat)))
tmp.range.of.days <-
cbind(
data = file.dat,
range_from_start_to_end_date_in_days = dat$range_from_start_to_end_date_in_days,
elapsed.dates.each.field = dat$elapsed.dates.each.field,
range_of_start_end_date = dat$range_of_start_end_date,
time.range.each.day = dat$time.range.each.day,
time.elapsed.each.day = dat$time.elapsed.each.day
)
range_table <- rbind(range_table, tmp.range.of.days)

tmp.one.hour.time.difference <- dat$one_hour_gap_time_stamp
tmp.one.hour.time.difference$field <- file.dat

if (dim(as.data.frame(tmp.one.hour.time.difference))[1] >= 2) {
one.hour.time.difference <-
rbind.fill(one.hour.time.difference, tmp.one.hour.time.difference)
} else{
one.hour.time.difference <- one.hour.time.difference
}


}

range_table <- as.data.frame(range_table)
range_table$days_number <-
as.numeric(gsub(
' days|day',
"",
range_table$range_from_start_to_end_date_in_days
))

# Range table
range_table
```
This is the range table for all fields. I have calculated the range for total
elapsed time (in terms of days) in `range_from_start_to_end_date_in_days`
column, total dates elapsed in `elapsed.dates.each.field` column, Range in dates
in `range_of_start_end_date` column, time range each day (from start to end
time) in `time.range.each.day.Time` column, total time spent in the field each
date in `time.elapsed.each.day.Time_elapsed (minutes)` column.

```{r}
# This is the list of consecutive rows with more than one hour time gap from all files:
one.hour.time.difference
```

I was able to find more than an hour gap in consecutive rows in `harvestC` data
only

```{r}
# Plotting the total number of days for each field type
ggplot(range_table, aes(x = data, y = days_number)) + geom_point(size = 4) +
  ggtitle("Plot in terms of elapsed time- \n start to end (in days)") +
theme_bw() +
theme(
axis.line = element_line(colour = "black"),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
panel.border = element_blank(),
panel.background = element_blank()
) +
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
theme(text = element_text(size = 14)) +
scale_y_continuous(breaks = round(c(
1,
max(range_table$days_number) * 1 / 3,
max(range_table$days_number) * 2 / 3,
max(range_table$days_number)
), 2)) +
# to check if any of the data were planted/harvested within one day range
geom_hline(yintercept = 1) +
geom_text((aes(3, 1, label = "Within 1 day (24 hours margin)", vjust = -1)))
```

Based on this plot, we can see that SeedA, seedB, seedC and seedD were all
planted within 24 hours span. We can also check the `range_table` above to get
the exact timeframe for this. However, if want to see if they were planted or
harvested on the same date, then we can plot this figure below:

```{r}
# If we want to see whether they were planted or harvested the same day (i.e
# Date) and (not necessarily 24 hours margin), we choose range_table$elapsed.dates.each.field


# Now plot them
ggplot(range_table, aes(x = data, y = elapsed.dates.each.field)) +
  geom_point(size = 2) +
  theme_bw() +
  ggtitle("Plot in terms of elapsed dates") +
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  theme(text = element_text(size=14)) +
# to check if any of the data were planted/harvested the same "Date"
geom_hline(yintercept = 1) +
  geom_hline(yintercept = 0, color="blue", linetype="dashed") +
geom_text((aes(3, 1, label = "Margin outside the \"same\" date", vjust = -1)))

```

This plot above shows that seedA and seedD were planted on the same date, and
harvestB was harvested also on the same date. Rest of the field data were
harvested/planted for multiple dates. Bubbles within the black solid line
indicates that the fields were, harvested/planted the same date. Only seedA,
seedD and harvestB are within the solid line. Here, Y axis margin of 1 means
next date. Data within the black and dashed blue lines indicate the same date. 


Additionally, if we want to see how many hours were spent on each field in
total, we can plot this below:

```{r}
plot.new <- range_table[, c("data", "time.elapsed.each.day.Time_elapsed (minutes)", "time.elapsed.each.day.Date")]
plot.new$time.elapsed.each.day.Time_elapsed <- as.numeric(plot.new$`time.elapsed.each.day.Time_elapsed (minutes)`)
plot.new$time.elapsed.each.day.Time_elapsed.in.hour <- plot.new$time.elapsed.each.day.Time_elapsed/60
plot.new.data <- setNames(aggregate(plot.new$time.elapsed.each.day.Time_elapsed.in.hour, list(range_table$data), sum), c("group", "time.spent.in.hours"))

ggplot(plot.new.data, aes(x = group, y = time.spent.in.hours)) +
  geom_point(size = 4) +
  theme_bw() +
  ggtitle("Plot in terms of total time spent in hours") +
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  theme(text = element_text(size=14)) 
```

Based on this above plot, it looks like only harvestC and harvestD had more than
10 hours spent on them.

Now, we can also plot start and end time for everyday record:
```{r}
range_table$start_date_time <-
as.POSIXct(paste(
range_table$time.range.each.day.Date ,
gsub("to.*| ", "", range_table$time.range.each.day.Time)
), tz = 'UTC')
range_table$end_date_time <-
as.POSIXct(paste(
range_table$time.range.each.day.Date ,
gsub(".*to| ", "", range_table$time.range.each.day.Time)
), tz = 'UTC')


tt <- range_table[, c("data", "start_date_time", "end_date_time")]

tt <- melt(tt, id = "data")



# Plot for date and time

ggplot(tt, aes(
x = data,
y = as.character.Date(value),
colour = variable
)) +
geom_point(size = 4) +
ylab("Dates and Time") +
theme_bw() +
ggtitle("Start and end time for everyday record") +
theme(
axis.line = element_line(colour = "black"),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
panel.border = element_blank(),
panel.background = element_blank()
) +
theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 14)) +
theme(axis.text.y = element_text(angle = 0, hjust = 1, size = 8)) +
# theme(text = element_text(size = 14)) +
geom_hline(yintercept = 1:length(tt$value), linetype = "dotted") +
labs(color = 'Each Day') +
scale_color_manual(labels = c("START_time", "END_time"),
values = c("blue", "red")) 

```
This plot also shows similar result as with previous plot, but with both start
and end time.

We can also check the moisture level differences in the fields by date and
times.
 
```{r}
# Now, let's work on the effect of time/dates on moiture level
tt1 <- harvestA
tt2 <- harvestB
tt3 <- harvestC
tt4 <- harvestD

tt1$type <- "harvestA"
tt2$type <- "harvestB"
tt3$type <- "harvestC"
tt4$type <- "harvestD"

make_posixct <- function(x) {
  x$Date <- gsub("T.*", "", x$Timestamp)
  x$Date <- as.factor(x$Date)
  x$Time <- gsub(".*T|Z", "", x$Timestamp)
  n <- 2
  pat <- paste0('^([^:]+(?::[^:]+){', n - 1, '}).*')
  x$Time  <- as.factor(sub(pat, '\\1', x$Time))
  return(x)
}

tt1 <- make_posixct(tt1)
tt2 <- make_posixct(tt2)
tt3 <- make_posixct(tt3)
tt4 <- make_posixct(tt4)
colnames(tt2) <- colnames(tt1)


df <- rbind(tt1, tt2, tt3, tt4)

df$Time <-
  as.POSIXct(strptime(df$Time, format = "%H:%M", tz = "UTC"))
lims <-
  as.POSIXct(strptime(c("0:00", "23:59"), format = "%H:%M", tz = "UTC"))
p <-
  ggplot(data = df, aes(
    x = Time,
    y = Moisture,
    colour = as.factor(Date)
  )) +
  geom_point() +
  scale_x_datetime(
    limits = lims,
    breaks = date_breaks("2 hour"),
    labels = date_format("%H:%M", tz = "UTC")
  ) +
  labs(color = 'Dates') +
  facet_grid(. ~ type)
p + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
```
Based on this figure above, we can see that the moisture content is very low
from midnight to 2:00pm, but then after, it increases substantially. I don't
think there is so much difference in moisture in terms of harvest dates though.


I tried to pull weather data for each observation, however the api I used only
allows 1000 requests per day, so I did not include that data.
```{r, eval= FALSE}
# We can use darksky like api's to pull weather data, however it's not
# possible as we have so many observations. Only allows 1000 downloads per day.

# install.packages("darksky")
library(darksky) darksky_api_key(force = TRUE)
#API: 60b8ca6aaece606b21c0dfec1cb7ec9f
head(seedA)
# field_or_seed <- seedA 
tmp <- get_forecast_for(10.25209, 478.4000, "2013-05-06T12:00:00")

# Create empty columns and cbind to the field data so you can add new weather
data field_or_seed <- cbind(field_or_seed, setNames(lapply(get.weather.cols,
function(x) x = NA), get.weather.cols)) 
field_or_seed$Date <- gsub("T.*", "", field_or_seed$Timestamp)
field_or_seed$Time <- gsub(".*T|Z", "", field_or_seed$Timestamp) # Range in days field_or_seed$date_time <-
as.POSIXct(paste(field_or_seed$Date, field_or_seed$Time), tz = "UTC")

field_or_seed_with_weather <- {}
for (i in 1:nrow(field_or_seed)) {
print(paste0("My iteration is: ", i)) tmp.min <-
get_forecast_for(44.311356, -96.798386, min(field_or_seed$date_time[i]))
tmp.weather.data <- cbind(field_or_seed[i,], tmp.min$daily)
field_or_seed_with_weather <- rbind(field_or_seed_with_weather,
tmp.weather.data)
}
```

Additional work:

Let's read our csv files from midterm

```{r}
# # If using local path
#   setwd(path.midterm)
#   all.files <- list.files(path.midterm)
  # here, instead, I am just using the URL
  all.files <- path.midterm
  all.files <- all.files[grepl("*.csv", all.files)]
  # > all.files
  # [1] "fieldA.csv" "fieldB.csv" "fieldC.csv" "fieldD.csv"
  

all.files <- all.files[grepl("fieldA.csv|fieldB.csv|fieldC.csv|fieldD.csv", all.files)]  
  
  
# reading all csv files within the path and saving them with their respective
# names
# https://raw.githubusercontent.com/achalneupane/data/master/
# reading all csv files within the path and saving them to their respective names.
for (i in 1:length(all.files)) {
assign(
gsub(".*master/|.csv", "", all.files)[i],
read.table(gsub("blob/", "",paste0("https://raw.githubusercontent.com",all.files[i])),
header = TRUE,
sep = ",")
)
print(paste0("Read file ", all.files[i], ' !'))
}
  
  # Now, let's merge field and seed data
  
  # head(fieldA)
  # head(seedA)
  
  dataA <-
  merge(
  fieldA,
  seedA,
  by.x = c("Easting", "Northing"),
  by.y = c("LonM", "LatM")
  )
  colnames(dataA) <- gsub(".x", "", colnames(dataA))
  dataB <-
  merge(
  fieldB,
  seedB,
  by.x = c("Easting", "Northing"),
  by.y = c("LonM", "LatM")
  )
  colnames(dataB) <- gsub(".x", "", colnames(dataB))
  dataC <-
  merge(
  fieldC,
  seedC,
  by.x = c("Easting", "Northing"),
  by.y = c("LonM", "LatM")
  )
  colnames(dataC) <- gsub(".x", "", colnames(dataC))
  dataD <-
  merge(
  fieldD,
  seedD,
  by.x = c("Easting", "Northing"),
  by.y = c("LonM", "LatM")
  )
  colnames(dataD) <- gsub(".x", "", colnames(dataD))
  
  # Taking 1000 ControlRates interval, we will also compare different dates (i.e,
  # ControlRate and date pairs) for each merged seed and field data.
  
  ControlRateInterval <- 1000
  # field A
  fieldA.data <-
  calculate_field_mean_SD_and_get_RR_EffSize(field = dataA, intervals = ControlRateInterval)
  # fieldA.data
  
  
  # field B
  fieldB.data <-
  calculate_field_mean_SD_and_get_RR_EffSize(field = dataB, intervals = ControlRateInterval)
  
  # fieldB.data
  
  # From midterm for fieldB, we saw significant difference between all Control Rates of
  # Seeding except between 28000 Vs 29000. The seeding rates of 28000 has the
  # highest Mean yield here.
  
  # field C
  fieldC.data <-
  calculate_field_mean_SD_and_get_RR_EffSize(field = dataC, intervals = ControlRateInterval)
  # fieldC.data
  
  # From midterm for fieldC, we saw all Control rates of seeding are significantly diffferent with the highest Mean yeild for 28000.
  
  # field D
  fieldD.data <-
  calculate_field_mean_SD_and_get_RR_EffSize(field = dataD, intervals = ControlRateInterval)
 # fieldD.data
```  
  
  Based on my analysis from midterm, we saw that it is better to compare all
  four fields together. Also, I do not want to clutter this report with too many
  figures and tables, I am analyzing 1000 Control Rate and all four fields
  merged together. This way we can compare the skewed data points if we analyze
  with date and ControlRate pairs. I have done this analysis below
  

  We can now merge all four fields  for SD, means and ControlRate level counts.

```{r}
  merged.SD.4.plots <-
  Reduce(
  function(x, y)
  merge(x, y, all = TRUE),
  list(
  fieldA.data$fieldSD,
  fieldB.data$fieldSD,
  fieldC.data$fieldSD,
  fieldD.data$fieldSD
  )
  )
  
  
  merged.Mean.4.plots <-
  Reduce(
  function(x, y)
  merge(x, y, all = TRUE),
  list(
  fieldA.data$field.mean,
  fieldB.data$field.mean,
  fieldC.data$field.mean,
  fieldD.data$field.mean
  )
  )
  
  
  merged.Count.4.plots <-
  Reduce(
  function(x, y)
  merge(x, y, all = TRUE),
  list(
  fieldA.data$field.Count,
  fieldB.data$field.Count,
  fieldC.data$field.Count,
  fieldD.data$field.Count
  )
  )
```

We need SD pooled for these levels so we can calculate Cohen's d and Required
Replicates.

```{r}  
  levels(merged.SD.4.plots$CR.Date.Levels)
  # First, we merged all merged.SD.4.plots, merged.Mean.4.plots and
  # merged.Count.4.plots from all fields.
  # Therefore, now merging all three dataframes from all four plots for SD, Mean
  # and counts by ControlRate column
  
  Mean_SD_Count.dat <- Reduce(function(...)
  merge(..., by = c("CR.Date.Levels", "grp"), all.x = TRUE),
  lapply(
  list(merged.Mean.4.plots, merged.SD.4.plots, merged.Count.4.plots),
  transform,
  grp = ave(seq_along(CR.Date.Levels), CR.Date.Levels, FUN = seq_along)
  ))
  # head(Mean_SD_Count.dat)
  
  # Now, we can also do Anova on the merged data `Mean_SD_Count.dat`
  
  Mean_SD_Count_merged_for_all_four_plots <- Mean_SD_Count.dat
  Mean_SD_Count_merged_for_all_four_plots$Yield <-
  Mean_SD_Count_merged_for_all_four_plots$Mean
  Mean_SD_Count_merged_for_all_four_plots$ControlRate.Levels <-
  Mean_SD_Count_merged_for_all_four_plots$ControlRate
  
  get_my_box_plot(Mean_SD_Count_merged_for_all_four_plots, plot_name = "all four fields")
```

  Based on this plot above, from all fields, we can tell that, Control Rates of
  seeding of 23000-25000 has significantly less Mean yield as compared to
  26000-29000. It looks like Control Rates of Seeding between 27000 and 28000
  has the highest Mean yield. I also think that there is no effect on seeding
  dates, I have done some analysis to show the seeding effect only in the last
  part of this report and based on that as well, I do not think there is 
  significant effect of Dates on Yield.

  
  We now calculate pooled SD for merged 4 plots `Mean_SD_Count.dat` to calculate
  Effect Size and Required Replicates for all fields (combined analysis) using
  pooled SD.
  
```{r}
pooled.dat <- Mean_SD_Count.dat


# # Pooled sd can be calculated as:
pooled.dat$df <- pooled.dat$Count - 1
# pooled SD is :
# pooledSD <- sqrt( sum(pooled.dat$sd^2 * pooled.dat$df) / sum(pooled.dat$df) )

# We can calculate our SD pooled using this formula:
#  s_{pooled} = \sqrt{\frac{\sum_i (n_i-1)s_i^2}{N-k}}
# We will derrive this in steps as below:

pooled.dat$df <- pooled.dat$Count - 1
pooled.dat$sd.square <- pooled.dat$SD ^ 2

pooled.dat$ss <- pooled.dat$sd.square * pooled.dat$df


# We can use convenience function (aggregate) for splitting and calculating
# the necessary sums.
ds <- aggregate(ss ~ CR.Date.Levels, data = pooled.dat, sum)

# Two different built-in methods for split apply, we could use aggregate for
# both if we wanted. This calculates our degrees of freedom.
ds$df <- tapply(pooled.dat$df, pooled.dat$CR.Date.Levels, sum)
# divide ss by df and then we get sd square
ds$sd.square <- ds$ss / ds$df
# Finally, we can get our sd pooled
ds$SD_pooled <- sqrt(ds$sd.square)
# ds

# However, we could also calculate our sd_pooled as below and get the same results :
# sd_pooled <- lapply( split(Mean_SD_Count.dat, Mean_SD_Count.dat$CR.Date.Levels),
# function(dd) sqrt( sum( dd$SD^2 * (dd$Count-1) )/(sum(dd$Count-1)-nrow(dd)) ) )

# Now, we calculate Mean (Mean of Means) from the merged table
# `Mean_SD_Count.dat so we can calculate Cohens d and RequiredReplicates for all
# four field combined.

ds.Mean <-
setNames(aggregate(
Mean_SD_Count.dat$Mean,
by = list(Mean_SD_Count.dat$CR.Date.Levels),
FUN = mean
),
c("CR.Date.Levels", "Mean"))

ds <- merge(ds, ds.Mean, by.x = "CR.Date.Levels")
# ds

# Now we calculate the Effect Size and Cohen's D for the combined 4 plots using
# mean yield and sd pooled for different ControlRate

# Now we calculate the Effect Size and Cohen's D for the combined 4 plots using
# mean yield and sd pooled for different ControlRate usinf our function

RequiredReplicates_for_all_fields <-
calculate_field_mean_SD_and_get_RR_EffSize(
field = ds,
intervals = ControlRateInterval,
Single.Field.Analysis = FALSE
)
```
This table gives us the required replicates and effectSize for all ControlRates-Date pairs
```{r}
RequiredReplicates_for_all_fields
```

We can plot this results

```{r}
RequiredReplicates_for_all_fields <-
  as.data.frame(RequiredReplicates_for_all_fields$Req.Rep.table.field)
  
  RequiredReplicates_for_all_fields$Group <-
  factor(RequiredReplicates_for_all_fields$Group)
  
  RequiredReplicates_for_all_fields$RequiredReplicates <-
  as.numeric(as.character(RequiredReplicates_for_all_fields$RequiredReplicates))
  
  RequiredReplicates_for_all_fields$EffectSize <-
  as.numeric(as.character(RequiredReplicates_for_all_fields$EffectSize))
  
  
  # There are one or two extreme Required replicates and extreme Effect size. We can
  # remove them to plot them, so we can properly visualize the datapoints
  
  RequiredReplicates_for_all_fields <-
  RequiredReplicates_for_all_fields[RequiredReplicates_for_all_fields$RequiredReplicates < 4000, ]
  RequiredReplicates_for_all_fields <-
  RequiredReplicates_for_all_fields[RequiredReplicates_for_all_fields$EffectSize < 1.5, ]
  
  my_levels <-
  length(levels(RequiredReplicates_for_all_fields$Group))
  RequiredReplicates_for_all_fields$EffectSize <-
  round(RequiredReplicates_for_all_fields$EffectSize, 2)
  
  # scaleFUN <- function(x) sprintf("%.2f", x)
  # Plot RequiredReplicates_for_all_fields
  
  n <- my_levels
  qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual', ]
  col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
  
  
  ggplot(
  RequiredReplicates_for_all_fields,
  aes(
  x = EffectSize,
  y = RequiredReplicates,
  shape = as.factor(Group),
  group = 1
  )
  ) +
  ggtitle("Effect size Vs Required replicates \n for all fields data \n(Combined Effect)") +
  geom_point(alpha = 0.9,
  size = 3,
  stroke = 1)  +
  scale_shape_manual(values = rep(c(0:2, 5:6, 9:10, 11:12, 14), times =
  4)) +
  scale_color_manual(values = col_vector) +
  geom_line(color = "black") +
  scale_x_continuous("EffectSize",
  breaks = c(0,
  0.2,
  0.5,
  1,
  max(
  as.numeric(RequiredReplicates_for_all_fields$EffectSize)
  ))) +
  theme_bw() +
  theme(
  axis.title = element_text(size = 14, face = "bold"),
  axis.text = element_text(size = 10),
  plot.title = element_text(hjust = 0.5, face = "bold"),
  axis.line = element_line(colour = "black"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank()
  ) +
  labs(shape = "Comparison groups")
```  
  
Based on this plot above, we can tell that there is strong negative correlation
between EffectSize and RequiredReplicates.


So far, we analyzed our data for controlRates-Date pair, now we can also do the
same analysis for Date effect only.
  
```{r}  
######################## Date Only Analysis !! ##############################

# If we want to do Date effect only without taking ControlRates into consideration, then we do the following:
  
# field A
fieldA.data <-
calculate_field_mean_SD_and_get_RR_EffSize(field = dataA,
intervals = ControlRateInterval,
Date.Only = TRUE)

# fieldA.data


# field B
fieldB.data <-
calculate_field_mean_SD_and_get_RR_EffSize(field = dataB,
intervals = ControlRateInterval,
Date.Only = TRUE)

# fieldB.data



# field C
fieldC.data <-
calculate_field_mean_SD_and_get_RR_EffSize(field = dataC,
intervals = ControlRateInterval,
Date.Only = TRUE)

# fieldC.data

# field D
fieldD.data <-
calculate_field_mean_SD_and_get_RR_EffSize(field = dataD,
intervals = ControlRateInterval,
Date.Only = TRUE)

# fieldD.data


# We can now merge all four fields  for SD, means and ControlRate level counts.

merged.SD.4.plots <-
Reduce(
function(x, y)
merge(x, y, all = TRUE),
list(
fieldA.data$fieldSD,
fieldB.data$fieldSD,
fieldC.data$fieldSD,
fieldD.data$fieldSD
)
)


merged.Mean.4.plots <-
Reduce(
function(x, y)
merge(x, y, all = TRUE),
list(
fieldA.data$field.mean,
fieldB.data$field.mean,
fieldC.data$field.mean,
fieldD.data$field.mean
)
)


merged.Count.4.plots <-
Reduce(
function(x, y)
merge(x, y, all = TRUE),
list(
fieldA.data$field.Count,
fieldB.data$field.Count,
fieldC.data$field.Count,
fieldD.data$field.Count
)
)

# We need SD pooled for these levels so we can calculate Cohen's d and
# Required Replicates.

levels(merged.SD.4.plots$CR.Date.Levels)

# First, we merged all merged.SD.4.plots, merged.Mean.4.plots and
# merged.Count.4.plots from all fields.

# Therefore, now merging all three dataframes from all four plots for SD, Mean
# and counts by ControlRate column

Mean_SD_Count.dat <- Reduce(function(...)
merge(..., by = c("CR.Date.Levels", "grp"), all.x = TRUE),
lapply(
list(merged.Mean.4.plots, merged.SD.4.plots, merged.Count.4.plots),
transform,
grp = ave(seq_along(CR.Date.Levels), CR.Date.Levels, FUN = seq_along)
))

head(Mean_SD_Count.dat)

# Now, we can also do Anova on the merged data `Mean_SD_Count.dat`

Mean_SD_Count_merged_for_all_four_plots <- Mean_SD_Count.dat
Mean_SD_Count_merged_for_all_four_plots$Yield <-
Mean_SD_Count_merged_for_all_four_plots$Mean
Mean_SD_Count_merged_for_all_four_plots$ControlRate.Levels <-
Mean_SD_Count_merged_for_all_four_plots$ControlRate

Mean_SD_Count_merged_for_all_four_plots$CR.Date.Levels <-
as.factor(gsub(
"-",
"_",
Mean_SD_Count_merged_for_all_four_plots$CR.Date.Levels
))


get_my_box_plot(Mean_SD_Count_merged_for_all_four_plots, plot_name = "all four fields")
```

  From this result above, since the residuals from ANOVA model has only 1 degree
  of freedom, and the global p-value is not significant, we don't have to
  plot the TUKEY pairwise test. Instead, we only show the Summary anova. This
  shows that the date alone doesn't affect our mean Yield.
  
  We now calculate pooled SD for merged 4 plots `Mean_SD_Count.dat` to calculate
  Effect Size and Required Replicates for all fields (combined analysis) using
  pooled SD.
  
```{r}
pooled.dat <- Mean_SD_Count.dat


# # Pooled sd can be calculated as:
pooled.dat$df <- pooled.dat$Count - 1
# pooled SD is :
# pooledSD <- sqrt( sum(pooled.dat$sd^2 * pooled.dat$df) / sum(pooled.dat$df) )


# We can calculate our SD pooled using this formula:
#  s_{pooled} = \sqrt{\frac{\sum_i (n_i-1)s_i^2}{N-k}}
# We will derrive this in steps as below:

pooled.dat$df <- pooled.dat$Count - 1
pooled.dat$sd.square <- pooled.dat$SD ^ 2

pooled.dat$ss <- pooled.dat$sd.square * pooled.dat$df


# We can use convenience function (aggregate) for splitting and calculating
# the necessary sums.
ds <- aggregate(ss ~ CR.Date.Levels, data = pooled.dat, sum)

# Two different built in methods for split apply, we could use aggregate for
# both if we wanted. This calculates our degrees of freedom.
ds$df <- tapply(pooled.dat$df, pooled.dat$CR.Date.Levels, sum)
# divide ss by df and then we get sd square
ds$sd.square <- ds$ss / ds$df
# Finally, we can get our sd pooled
ds$SD_pooled <- sqrt(ds$sd.square)
ds

# However, we could also calculate our sd_pooled as below and get the same results :
sd_pooled <-
lapply(split(Mean_SD_Count.dat, Mean_SD_Count.dat$CR.Date.Levels),
function(dd)
sqrt(sum(dd$SD ^ 2 * (dd$Count - 1)) / (sum(dd$Count - 1) - nrow(dd))))

# Now, we calculate Mean (Mean of Means) from the merged table
# `Mean_SD_Count.dat`, so we can calculate Cohens d and RequiredReplicates for
# all four field combined.

ds.Mean <-
setNames(aggregate(
Mean_SD_Count.dat$Mean,
by = list(Mean_SD_Count.dat$CR.Date.Levels),
FUN = mean
),
c("CR.Date.Levels", "Mean"))

ds <- merge(ds, ds.Mean, by.x = "CR.Date.Levels")
ds

# Now we calculate the Effect Size and Cohen's D for the combined 4 plots using
# mean yield and sd pooled for different ControlRate

# Now we calculate the Effect Size and Cohen's D for the combined 4 plots using
# mean yield and sd pooled for different ControlRate usinf our function

# `calculate_field_mean_SD_and_get_RR_EffSize`.

RequiredReplicates_for_all_fields <-
calculate_field_mean_SD_and_get_RR_EffSize(
field = ds,
intervals = ControlRateInterval,
Single.Field.Analysis = FALSE
)

RequiredReplicates_for_all_fields

# We can plot this results for better visualization of pattern.

RequiredReplicates_for_all_fields <-
as.data.frame(RequiredReplicates_for_all_fields$Req.Rep.table.field)

RequiredReplicates_for_all_fields$Group <-
factor(RequiredReplicates_for_all_fields$Group)

RequiredReplicates_for_all_fields$RequiredReplicates <-
as.numeric(as.character(RequiredReplicates_for_all_fields$RequiredReplicates))

RequiredReplicates_for_all_fields$EffectSize <-
as.numeric(as.character(RequiredReplicates_for_all_fields$EffectSize))


# There are one or two extreme Required replicates and extreme Effect size. We can
# remove them to plot them, so we can properly visualize the datapoints

RequiredReplicates_for_all_fields <-
RequiredReplicates_for_all_fields[RequiredReplicates_for_all_fields$RequiredReplicates < 4000, ]

RequiredReplicates_for_all_fields <-
RequiredReplicates_for_all_fields[RequiredReplicates_for_all_fields$EffectSize < 1.5, ]



my_levels <-
length(levels(RequiredReplicates_for_all_fields$Group))
RequiredReplicates_for_all_fields$EffectSize <-
round(RequiredReplicates_for_all_fields$EffectSize, 2)

# Plot RequiredReplicates_for_all_fields
n <- my_levels
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))


ggplot(
RequiredReplicates_for_all_fields,
aes(
x = EffectSize,
y = RequiredReplicates,
shape = as.factor(Group),
group = 1
)
) +
ggtitle("Effect size Vs Required replicates for all fields data \n(Combined Effect)") +
geom_point(alpha = 0.9,
size = 3,
stroke = 1)  +
scale_shape_manual(values = rep(c(0:2, 5:6, 9:10, 11:12, 14), times =
4)) +
scale_color_manual(values = col_vector) +
geom_line(color = "black") +
scale_x_continuous("EffectSize",
breaks = c(0,
0.2,
0.5,
1,
max(
as.numeric(RequiredReplicates_for_all_fields$EffectSize)
))) +
theme_bw() +
theme(
axis.title = element_text(size = 14, face = "bold"),
axis.text = element_text(size = 10),
plot.title = element_text(hjust = 0.5, face = "bold"),
axis.line = element_line(colour = "black"),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
panel.border = element_blank(),
panel.background = element_blank()
) +
labs(shape = "Comparison groups")

################################ End of codes !! ################################
```

Discussion: Based on these results, I think that the fields were planted within
24 hours margin. seedB and seedC were planted on the different dates.
Similarly, harvestB was harvested within 24 hours and on the same date, but rest
of the fields were harvested/planted for multiple dates. Additionally, I also
performed ANOVA for ControlRates paired with dates and found no significant
difference to  what I found for my midterm. I analyzed these data for Date
effect also and it looks like there is no Date effect on Mean Yield. I thought
the skewned in 28000 interval from my midterm would be addressed by Date effect,
but I could not find significant effect of date here. As in the mid-term, I also
calculated the required replicated for ControlRate-Date pairs as well as Date
only observations, and in both cases, the EffectSize was inversely proportional
to RequiredReplicates.

