---
Authors: ["**Achal Neupane**"]
title: "Logistic Regression&GLM-I"
date: 2019-10-01T17:26:23-05:00
draft: false
output: html_document
tags:
- R
- Statistics
- Modern Applied Statistics I
summary: Statistics series
---

<script type="text/javascript" src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.1/MathJax.js?config=TeX-AMS-MML_HTMLorMML">
</script>


Answer all questions specified on the problem and include a discussion on how
your results answered/addressed the question.

Submit your \textbf{.rmd} file with the knitted \textbf{PDF} (or knitted Word
Document saved as a PDF). If you are having trouble with .rmd, let us know and
we will help you, but both the .rmd and the PDF are required.

This file can be used as a skeleton document for your code/write up. Please
follow the instructions found under Content for Formatting and Guidelines. No
code should be in your PDF write-up unless stated otherwise.

For any question asking for plots/graphs, please do as the question asks as well
as do the same but using the respective commands in the GGPLOT2 library. (So if
the question asks for one plot, your results should have two plots. One produced
using the given R-function and one produced from the GGPLOT2 equivalent). This
doesn't apply to questions that don't specifically ask for a plot, however I
still would encourage you to produce both.

You do not need to include the above statements.

Please do the following problems from the text book R Handbook and stated.

1. Collett (2003) argues that two outliers need to be removed from the
\textbf{plasma} data. Try to identify those two unusual observations by means of
a scatterplot. (7.2 on Handbook)

```{r}
library("calibrate")
library("ggplot2")
library("HSAUR3")
library("knitr")
data("plasma")

plasma$rownumber <- 1:nrow(plasma)
plot.new()
# square plotting region, independent of device size; 1 x 1 pictures on one plot
par(mfrow = c(1, 1), pty = "s")
plot(
  plasma$fibrinogen,
  plasma$globulin,
  col = plasma$ESR,
  data = plasma,
  pch = 18,
  main = "base R: Scatterplot of plasma data",
  xlab = "Fibrinogen",
  ylab = "Globulin"
)
text(globulin ~fibrinogen, labels = rownumber,data=plasma, cex=0.8, font=0.5)
abline(v = 3.5)
abline(h = 45)
# points(plasma[c(27, 30, 32), 1:2], pch = 5)
legend(
  "bottomright",
  c("ESR<20", "ESR>20"),
  # title = "ESR",
  inset = c(0, 1),
  xpd=TRUE, 
  horiz=TRUE,
  col = c("black", "red"),
  # lty = c(1, 1),
  pch = c(18, 18),
  bty = "n"
)


ggplot(data = plasma, aes(x = fibrinogen, y = globulin, colour = ESR)) +
  geom_point() +
  ggtitle("ggplot: Scatterplot of Plasma data") +
  xlab("Fibrinogen") +
  ylab("Globulin") +
  geom_text(aes(label=plasma$rownumber),hjust=0, vjust=0) +
  geom_hline(yintercept=45, linetype="dashed", 
             color = "black", size=0.5) +
  geom_vline(xintercept=3.5, linetype="dashed", 
             color = "black", size=0.5)

```

Discussion: Based on the scatter plot (and if we consider all data points, both ESR > and < 20, together), we could say that rows 27, 30 and 32 as the potential outliers. If we separate the data by ESR >20 and ESR < 20, there may be other outliers for each group. I think we could make box plots by group to determine the outliers more effectively.



2. (Multiple Regression) Continuing from the lecture on the \textbf{hubble} data
from \textbf{gamair} library;

    a) Fit a quadratic regression model, i.e.,a model of the form
$$\text{Model 2:   } velocity = \beta_1 \times distance + \beta_2 \times distance^2 +\epsilon$$

```{r}
library(gamair)
data(hubble)

# A. Fit a quadratic regression model
model2 <- lm(y~x + I(x^2) -1, data = hubble)
# summary(model2)
kable(summary(model2)$coefficients, caption = "Summary of the model2")
```
   
    b) Plot the fitted curve from Model 2 on the scatterplot of the data
    
```{r}
# fitted curve
# index  <- seq(0, 22, 0.1)
index <- seq(min(hubble$x),max(hubble$x),0.1)
index2 <- index^2
# predicted <- predict(model2,list(x = index, x2=index2))
predicted <- model2$fitted.values
#create a data frame of x nd y values for plotting for ggplot
data <- as.data.frame(cbind(x = hubble$x,predicted))
# Scatter Plot
plot.new()
par(mfrow = c(1, 1), pty = "s")
plot(y~x, data = hubble, main = "base R: Scatter plot with fitted curve from Model2", xlab = "Distance", ylab = "Velocity")
lines(data$x[order(data$x)], data$predicted[order(data$predicted)], col = "green")

ggplot(data = model2, aes(x = model2$model$x, y = model2$model$y)) +
  geom_point() +
  geom_line(aes(x = model2$model$x, y = model2$fitted.values), colour = "green") +
  labs(title = "ggplot: Scatter plot with fitted curve from Model2", x = "Distance", y = "velocity")

```
    
    c) Add the simple linear regression fit (fitted in class) on this plot - use
    different color and line type to differentiate the two and add a legend to
    your plot.
    
```{r}    
# Simple lm fitted in class
hmod <- lm(y~x - 1 , data = hubble)
plot.new()
par(mfrow = c(1, 1), pty = "s")
plot(y~x, data = hubble, main = "base R: scatter plot for hubble data", xlab = "Distance", ylab = "Velocity")
lines(data$x[order(data$x)], data$predicted[order(data$predicted)], col = "green")
abline(hmod, lty=2, col=2)

# Legend
legend("bottomright", c("Quadratic", "Linear"), lty = 1:2, col = 2:1)


## ggplot version

ggplot(data = model2, aes(x = model2$model$x, y = model2$model$y)) +
  geom_point() +
  geom_line(aes(x = model2$model$x, y = model2$fitted.values, colour = "Quadratic")) +
  geom_line(data = hmod, aes(x = hmod$model$x, y = hmod$fitted.values, colour = "Linear")) +
  labs(title = "ggplot: Scatter plot with fitted curve from Model2", x = "Distance", y = "velocity", colour = "Models")
```
    
    d) Which model do you consider most sensible considering the nature of the data - looking at the plot? 
    
    Answer: The simple model seems more sensible to me.  The data points seem to
    follow a line from lower left to upper right of the plot without a clear
    curvature. However, strictly saying, there isn't much difference between the two models.
    
    e) Which model is better? - provide a statistic to support you claim.
    
    Note: The quadratic model here is still regarded as a ``linear regression"
    model since the term ``linear" relates to the parameters of the model and
    not to the powers of the explanatory variables.
    
```{r}
summary(model2) # # Quadratic regression model
mod.2 <- summary(model2)
summary(hmod)  # Simple linear model
hmod.1 <- summary(hmod) 
cat ("Adjusted R-square")
kable(cbind(Quadratic = mod.2$adj.r.squared, Linear = hmod.1$adj.r.squared), caption = "Adjusted R-square", row.names = FALSE )

```

    Answer to 2e. The statistics appear to support the simple model as the
    better one. Since the Adjusted r-squared statistic is higher for the simple model
    (0.9394) Vs. Quadratic (0.9388554) which indicates that the simple model explains more of the variability in the response data than does the quadratic model.


3. The \textbf{leuk} data from package \textbf{MASS} shows the survival times
from diagnosis of patients suffering from leukemia and the values of two
explanatory variables, the white blood cell count (wbc) and the presence or
absence of a morphological characteristic of the white blood cells (ag).

    a) Define a binary outcome variable according to whether or not patients
    lived for at least 24 weeks after diagnosis. Call it \textit{surv24}.
    
```{r}
#add a binary column named surv24 for time greater than or less than 24. 
library(MASS)
library(dplyr)
q3_subset <- leuk %>%
  mutate(surv24 = ifelse(time >= 24, 1,0))
```

    b) Fit a logistic regression model to the data with \textit{surv24} as
    response. It is advisable to transform the very large white blood counts to
    avoid regression coefficients very close to 0 (and odds ratio close to 1).
    You may use log transformation.

```{r}
surv24.model <- glm(surv24 ~ log(wbc) + ag, data=q3_subset,family = 'binomial')
kable(summary(surv24.model)$coefficient, caption = "Summary coefficients of the glm")
```

    c) Construct some graphics useful in the interpretation of the final model you fit. 
    
```{r}
cat("Create a scatter plot of the data fitting the two curves of test results to the fitted output of the model prediciton")

x.extension <- seq(0, max(log(q3_subset$wbc)+4.5), by = 0.5)
espframe <- data.frame("x.extension" = x.extension, "agpress" = (exp(surv24.model$coefficients[1] +surv24.model$coefficients[2]*x.extension + surv24.model$coefficients[3])/(1+exp(surv24.model$coefficient[1] + surv24.model$coefficients[2]*x.extension + surv24.model$coefficients[3]))), "agabs" = exp(surv24.model$coefficients[1] +surv24.model$coefficients[2]*x.extension)/(1+exp(surv24.model$coefficient[1] + surv24.model$coefficients[2]*x.extension)))


plot.new()
par(mfrow = c(1, 1), pty = "s")
plot(x = log(leuk$wbc), y = surv24.model$fitted.values, col = leuk$ag, xlim = c(0,15), ylim = c(0,1), ylab = "Survive (Time, surv24wks)", xlab = "log (wbc counts)", main = "base R: plot of logistic model of Leuk data")
lines(x = x.extension, y = exp(surv24.model$coefficients[1] +surv24.model$coefficients[2]*x.extension)/(1+exp(surv24.model$coefficient[1] + surv24.model$coefficients[2]*x.extension)))
lines(x = x.extension, y = exp(surv24.model$coefficients[1] +surv24.model$coefficients[2]*x.extension + surv24.model$coefficients[3])/(1+exp(surv24.model$coefficient[1] + surv24.model$coefficients[2]*x.extension + surv24.model$coefficients[3])))
legend("bottomleft", legend = c("Ag Absent", "Ag Present"), col = c("black", "red"), lty = c(1,1))

leuk.gg <- data.frame("logwbc" = log(leuk$wbc), surv24 = q3_subset$surv24, "fv" = surv24.model$fitted.values, "ag" = leuk$ag)

leuk.gg <- cbind(leuk.gg, espframe)
ggplot(leuk.gg, aes(x = logwbc, y = fv, colour = ag)) + 
  geom_point() +
  # scale_colour_discrete(guide = FALSE) +
  # guides(colour=FALSE) +
  geom_line(aes(x = x.extension, y = agpress, colour = "present")) +
  geom_line(aes(x = x.extension, y = agabs, colour = "absent")) +
  labs ( title = "ggplot: plot of logistic model of Leuk data", x = "log of WBC count", y = "Survive (Time, surv24wks)")




cat("Survival Vs WBC count with logistic model on actual data points")

# # base plot version
line.1.dat <- leuk.gg[leuk.gg$ag == 'absent', ]
line.2.dat <- leuk.gg[leuk.gg$ag == 'present', ]
plot.new()
par(mfrow = c(1, 1), pty = "s")
plot(
  x = leuk.gg$logwbc,
  y = leuk.gg$surv24,
  xlim=c(0,15),
  ylim = c(0,1),
  col = leuk.gg$ag,
  xlab = "WBC counts",
  ylab = "Probability of Death prior to 24 Weeks",
  main = "base R: Survival Vs WBC Counts in Leukaemia Patients"
)
lines(x.extension, leuk.gg$agpress, col = "green")
lines(x.extension, leuk.gg$agabs, col = "black")
legend(
  "topleft",
  title = "AG test",
  legend = c("absent", "present"),
  inset = c(1, 0),
  xpd = TRUE,
  horiz = FALSE,
  col = c("black", "green"),
  lty = c(1,1),
  pch = c(1, 2),
  bty = "n"
)

ggplot(leuk.gg, aes(x = logwbc, y = surv24, color = ag)) +
  geom_point() +
  scale_colour_manual(name = "AG test", values = c('black', 'green')) +
  geom_line(aes(x = x.extension, y = agpress, colour = "present")) +
  geom_line(aes(x = x.extension, y = agabs, colour = "absent")) +
  labs(title = 'ggplot: Survival Vs WBC Counts in Leukaemia Patients',
       x = 'log WBC Count',
       y = 'Probability of Death prior to 24 Weeks') +
  theme_classic()
```
    
    d) Fit a model with an interaction term between the two predictors. Which
    model fits the data better? Justify your answer.
    
```{r}

#fitting the model with the interaction term ag * log(wbc)
surv24.model2 <- lm(surv24 ~ ag * log(wbc), data=q3_subset,family='binomial')
kable(summary(surv24.model2)$coefficients, caption = "Summary of the linear model with an interaction")

mod2 = summary(surv24.model2)
mod = summary(surv24.model)
# we can also calculate adjusted r-square value for glm using 
library(rsq)
mod.rsq.adj = rsq(surv24.model,adj=TRUE,type=c('v','kl','sse','lr','n'),data=NULL)
mod2.rsq.adj = rsq(surv24.model2,adj=TRUE,type=c('v','kl','sse','lr','n'),data=NULL)
# if not using package rsq
# adj.rsq = rbind(mod2$adj.r.squared, (1 -(mod$deviance/mod$null.deviance)) * 32/(32-2-2))
adj.rsq = rbind(mod2.rsq.adj, mod.rsq.adj)

row.names(adj.rsq) <- c("Linear model with interation", "Linear model")
kable(adj.rsq, col.names = "Adjusted R-square values")
```

Since the adjusted R-square value for Linear model with the interacion is higher, I would say the model with an interaction fits the data better.


4. Load the \textbf{Default} dataset from \textbf{ISLR} library. The dataset
contains information on ten thousand customers. The aim here is to predict which
customers will default on their credit card debt. It is a four-dimensional
dataset with 10000 observations. The question of interest is to predict
individuals who will default . We want to examine how each predictor variable is
related to the response (default). Do the following on this dataset

    a) Perform descriptive analysis on the dataset to have an insight. Use
    summaries and appropriate exploratory graphics to answer the question of
    interest.
    
    b) Use R to build a logistic regression model. 
    
    c) Discuss your result. Which predictor variables were important? Are there
    interactions?
    
    d) How good is your model? Assess the performance of the logistic regression
    classifier. What is the error rate?
  
```{r}

# Set up data
data("Default", package = "ISLR")

kable(summary(Default[,1:2]), caption = "Summary of default and student status")
kable(summary(Default[,3:4]), caption = "Summary of Income and Balance")

#create default binary
default_binary     <-
  ifelse(regexpr('Yes', Default$default) == -1, 0, 1)
dflt_str <-
  ifelse(regexpr('Yes', Default$default) == -1,
         "Not Defaulted",
         "Defaulted")

stdn     <- ifelse(regexpr('Yes', Default$student) == -1, 0, 1)
stdn_str <-
  ifelse(regexpr('Yes', Default$student) == -1, "Not-Student", "Student")

blnc <- Default$balance
incm <- Default$income

df <-  data.frame(default_binary, dflt_str, stdn, stdn_str, blnc, incm)

# par(mfrow = c(1, 1))

cat("Balance appears roughly normal")
hist(blnc, main = "Balance")
# ggplot() + geom_histogram(aes(blnc), bins = 13, color = "black", fill = "white")


cat("Income appears roughly normal with two means")
hist(incm, main = "Income")

cat("Dual means in income appears explained by student status")
layout(matrix(1:2, ncol = 2))
hist(
  subset(df$incm, df$stdn == 1),
  main = "Income by Student Status",
  ylab = "Income",
  xlab = "Student: Yes"
)
hist(
  subset(df$incm, df$stdn == 0),
  main = "",
  ylab = "Income",
  xlab = "Student: No"
)

cat("**And** the dual means in income appears NOT to be explained by default status")
layout(matrix(1:2, ncol = 2))
hist(
  subset(df$incm, df$default_binary == 1),
  main = "Income by Default Status",
  ylab = "Income",
  xlab = "Default: Yes"
)
hist(
  subset(df$incm, df$default_binary == 0),
  main = "",
  ylab = "Income",
  xlab = "Default: No"
)

cat("Clustering of income v. balance explained by student status")
plot.new()
par(mfrow = c(1, 1), pty = "s")
plot(
  Default$income ~ Default$balance,
  col = Default$student,
  main = "base R: Income by Balance",
  ylab = "Income",
  xlab = "Balance",
  pch = 18
)
legend(
  "topright",
  c("Yes", "No"),
  title = "Student?",
  # bty = "n",
  fill = c("red", "black"),
  pch = c(18,18)
)

ggplot(data = Default, aes(x = balance, y = income, colour = student)) + 
  geom_point() +
  labs(title = "ggplot: Income by Balance") + 
  guides(colour=guide_legend(title="Student?")) +
  scale_color_manual(values = c("No" = "black", "Yes" = "red"))

plot.new()
par(mfrow = c(1, 1), pty = "s")
boxplot(balance~student, data = Default, main = "base R: Balance grouped by Student status", xlab = "student", ylab = "balance")

ggplot(data = Default, aes(x = student, y = balance)) +
  geom_boxplot() +
  labs(title = "ggplot: Balance grouped by Student status")

plot.new()
par(mfrow = c(1, 1), pty = "s")
boxplot(balance~default, data = Default, main = "base R: Balance grouped by Default status")

ggplot(data = Default, aes(x = default, y = balance)) +
  geom_boxplot() +
  labs(title = "ggplot: Balance grouped by Default status")

plot.new()
par(mfrow = c(1, 1), pty = "s")
boxplot(income~student, data = Default, main = "base R: Income grouped by Student status")

ggplot(data = Default, aes(x = student, y = income)) +
  geom_boxplot() +
  labs(title = "ggplot: Income grouped by Student status")

plot.new()
par(mfrow = c(1, 1), pty = "s")
boxplot(income~default, data = Default, main = "base R: Income grouped by Default status")

ggplot(data = Default, aes(x = default, y = income)) +
  geom_boxplot() +
  labs(title = "ggplot: Income grouped by Default status")


cat("Median and Max income are lower for defaulted than not defaulted loans")
tapply(df$incm, df$dflt_str, FUN = summary)

cat("Median and max balance are higher for defaulted rather than not defaulted loans")
tapply(df$blnc, df$dflt_str, FUN = summary)

cat("#B. Use R to build a logistic regression model")

# # https://stackoverflow.com/questions/13366755/what-does-the-r-formula-y1-mean
regression_model0 <- glm(default_binary ~ stdn + blnc + incm, family = binomial())
summary(regression_model0)

# # we could also do to select all predictors in the data
# mod <- glm(default~., data = Default, family = binomial)
# summary(mod)

cat("Then with interactions:")

regression_model1 <- glm(default_binary ~ stdn + blnc + incm + stdn * blnc + stdn * incm + blnc * incm, family = binomial())
summary(regression_model1)



cat("# D. Error Rate")
dflt.fitted0 <- predict(regression_model0, type = "response")
dflt.fitted1 <- predict(regression_model1, type = "response")

levs <- c("Defaulted", "Not Defaulted")
Tr <- default_binary

Predicted0 <-
  factor(ifelse(dflt.fitted0 >= 0.50, "Defaulted", "Not Defaulted"),
         levels = levs)
Predicted1 <-
  factor(ifelse(dflt.fitted1 >= 0.50, "Defaulted", "Not Defaulted"),
         levels = levs)
Tr1 <-
  factor(ifelse(Tr >= 0.50, "Defaulted", "Not Defaulted"), levels = levs)
rate0 <- table(Predicted0, True = Tr1)
rate1 <- table(Predicted1, True = Tr1)
rate0
error_rate0 <- 1 - (rate0[1, 1] + rate0[2, 2]) / sum(rate0)
error_rate0

rate1
error_rate1 <- 1 - (rate1[1, 1] + rate1[2, 2]) / sum(rate1)
error_rate1

cat("analysis of variance")
anova(regression_model0, regression_model1, test = 'Chisq')

```

4a. Based on the outputs for 4a. we can tell that:

* Fewer people default than don’t default.

* Defaulters and non-defaulters appear to have the same income range, given student status.

* Defaulters appear to have higher balances.

* If students default, they likely do it with over $1,000 balance.

*	If non-students default, they are likely do it with over $500 balance.

4c. 
Without taking interactions into account, it appears that two predictors-
student and balance are significant. With interactions involved, it appears that
only balance predictor is important.

4d.
The model without interactions has an AICof 1579.5 and the interaction model has
an AIC of 1585.1 (slightly higher). But, both have almost similar error rates
~2.7 %. Also, since analysis of deviance also shows that the chi-square test has
no significance at 5% level, we can conclude that both models are almost the
same as a working model.

5. Go through Section 7.3.1 of the Handbook. Run all the codes (additional exploration of data is allowed) and write your own version of explanation and interpretation.

```{r}
cat("# density plot")
plasma <- plasma

layout(matrix(1:2,ncol=2))
cdplot(ESR ~ fibrinogen, data=plasma)
cdplot(ESR ~ globulin,data=plasma)
```

It appears that above a certain level of fibrogen, ESR drops sucessively. This is not the case for globulin.
 
ESR Logistic Regression an Confidence Interval Estimates:

```{r}
plasma_glm_1 <- glm(ESR ~ fibrinogen, data = plasma, family=binomial())
confint(plasma_glm_1,parm='fibrinogen')
```

Here, fibrinogen might have value as a predictor of ESR. We can look at the summary.

```{r}
summary(plasma_glm_1)
```

The summary output indicates a 5% significance of fibrinogenand and increase of the log-odds of ESR > 20 by about 1.83 with confidence interval (CI) of 0.33 to 3.99.

```{r}
exp(coef(plasma_glm_1)['fibrinogen'])
```
 
Fibrinogen might have value as a predictor of ESR. 
To make the results more readable, it is useful to apply an exponent function. This exponenetiates the log-odds of fibriogen and CI to correspond with the data.

```{r}
exp(confint(plasma_glm_1, parm='fibrinogen'))
```

We can also perform logistic regression of both explanatory variables (fibrinogen and globulin) and text for the deviance.

```{r}
plasma_glm_2 <- glm(ESR ~ fibrinogen + globulin, data = plasma, family = binomial())
summary(plasma_glm_2)

cat("# comparison of models")
anova(plasma_glm_1, plasma_glm_2, test= 'Chisq')

```

Now, we can make the bubble plot of the predicted values of model II (plasma_glm_2). The plot shows that the probablity of ‘good’ ESR reading increases as fibrinogen increases. This is true of globulin only up to a point.

```{r}
prob <- predict(plasma_glm_2, type='response')

plot.new()
par(mfrow = c(1, 1), pty = "s")
plot(globulin ~ fibrinogen,data=plasma,xlim=c(2,6),ylim=c(25,55),pch='.', main = "Bubble plot of the predicted values of model II")
symbols(plasma$fibrinogen,plasma$globulin,circles=prob,add=T)
```