---
Authors: ["**Achal Neupane**"]
title: "Quantile Regression"
date: 2019-10-18T17:26:23-05:00
draft: false
output: html_document
tags:
- R
- Statistics
- Modern Applied Statistics I
summary: Statistics series
---


<script type="text/javascript" src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.1/MathJax.js?config=TeX-AMS-MML_HTMLorMML">
</script>


Answer all questions specified on the problem and include a discussion on how your results answered/addressed the question.

Submit your \textbf{.rmd} file with the knitted \textbf{PDF} (or knitted Word Document saved as a PDF). If you are having trouble with .rmd, let us know and we will help you, but both the .rmd and the PDF are required.

This file can be used as a skeleton document for your code/write up. Please follow the instructions found under Content for Formatting and Guidelines. No code should be in your PDF write-up unless stated otherwise.

For any question asking for plots/graphs, please do as the question asks as well as do the same but using the respective commands in the GGPLOT2 library. (So if the question asks for one plot, your results should have two plots. One produced using the given R-function and one produced from the GGPLOT2 equivalent). This doesn't apply to questions that don't specifically ask for a plot, however I still would encourage you to produce both.

You do not need to include the above statements.

Please do the following problems from the text book R Handbook and stated.

1. Consider the {\textbf{clouds}} data from the {\textbf{HSAUR3}} package
   
    a) Review the linear model fitted to this data in Chapter 6 of the text book and report the model and findings. 
      
    b) Fit a median regression model. 
      
    c) Compare the two results. 
    
```{r}
library(tidyverse)
library(gridExtra)
library(HSAUR3)
library(mboost)
library("quantreg")
library("rpart")
library("TH.data")
library(gamlss.data)
library(lattice)
data(clouds)
head(clouds)

cat("1a. Review the linear model fitted to this data in Chapter 6")
clouds_formula <- rainfall ~ seeding + seeding:(sne + cloudcover + prewetness + echomotion) + time
clouds.lm <- glm(clouds_formula, data = clouds)
summary(clouds.lm)

cat("It looks like the seedingyes variable is the most significant variable in the model followed by seedingyes:sne")

cat("Now we choose continous variable sne to fit our linear model")
clouds.lm <- glm(rainfall ~ sne, data = clouds)
summary(clouds.lm)
MSE.lm <- mean((predict(clouds.lm, data = clouds)-clouds$rainfall)^2)
cat("Linear model MSE: ",MSE.lm)
ggplot(data=clouds, aes(x=sne, y=rainfall, col=seeding)) + geom_point() + geom_smooth(method='lm') + labs(title='Rainfall determined by suitability criterion',x='S-NE Criterion', y='Rainfall')


cat("1b. Fit a median regression (quantile regression) model")
median.reg <- rq(rainfall ~ sne, data = clouds, tau = 0.5)
cat("Summary of the model")
summary(median.reg)
cat("Summary of model with bootstrapped standard error")
summary(median.reg, se = "boot" )
MSE.mrm <- mean((predict(median.reg, data = clouds, type = "response")-clouds$rainfall)^2)
cat("MSE of median regression Model: ",MSE.mrm)

cat("1c. We can also plot this model to compare with previous linear model")

plot_linear <- ggplot(data=clouds, aes(x=sne, y=rainfall, col=seeding)) + geom_point() + geom_smooth(method='lm',se=FALSE) + labs(title='Linear regression model: Rainfall\n determined by suitability criterion',x='S-NE Criterion', y='Rainfall') + theme_classic()


plot_median <- ggplot(data=clouds, aes(x=sne, y=rainfall, col=seeding)) + geom_point()  + labs(title='Median regression model: Rainfall\n determined by suitability criterion',x='S-NE Criterion', y='Rainfall') + stat_quantile(quantiles=c(0.50), method='rq') + theme_classic()

grid.arrange(plot_linear, plot_median, ncol=2)

```
    

Discussion:

In addition to fitting the models asked in 1a and 1b, I used graphical methods to compare a linear regression fit vs median regression fit for 1c. As shown in the plots, a median regression model on the data splits the absence of seeding in such a way that the median regression line has a positive slope, whereas, the  simple linear regression model seems to have the negative slope. This indicates that there is higher variability in the rainfall data when cloud seeding is absent. Additionally, in presence of seeding, the median regression line is not weighted by the outliers, therefore seems better at explaining the overall data due to the high variability of the rainfall variable.



2. Reanalyze the {\textbf{bodyfat}} data from the {\textbf{TH.data}} package. 

    a) Compare the regression tree approach from chapter 9 of the textbook to median regression and summarize the different findings.
    
    b) Choose one independent variable. For the relationship between this variable and DEXfat, create linear regression quantile models for the 25%, 50% and 75% quantiles. Plot DEXfat vs that independent variable and plot the lines from the models on the graph.
    
    
```{r}
data("bodyfat")
head(bodyfat)
ncol(bodyfat)
nrow(bodyfat)
library("rpart")
# ?bodyfat

cat("First, fit regression tree model")
#bodyfat_formula <- DEXfat ~ age + waistcirc + hipcirc + elbowbreadth + kneebreadth
bodyfat_rp  <- rpart(DEXfat ~ age + waistcirc + hipcirc + elbowbreadth + kneebreadth, data = bodyfat, control = rpart.control(minsplit=10))
summary(bodyfat_rp)

cat("Plot the regression tree model")
library(partykit)
plot(as.party(bodyfat_rp), tp_args = list(id=FALSE), main = "Regression tree of the model")

cat("Print CP-table")
print(bodyfat_rp$cptable)

cat("CP value with lowest xerror")
min.cp <- which.min(bodyfat_rp$cptable[,'xerror'])
min.cp

cat(
  "Following the methods in the book, we can fit the model\n using lowest xerror rate from CP for prunning tree model"
)

#extract the lowest CP
cp <- bodyfat_rp$cptable[min.cp, 'CP']
bodyfat_prune <- prune(bodyfat_rp, cp=cp)

cat("summary of the median regression model")
#Median Regression model
bodyfat_rpart_qrm <- rq(DEXfat ~ age + waistcirc + hipcirc + elbowbreadth + kneebreadth, data=bodyfat, tau = 0.50)
summary(bodyfat_rpart_qrm)
bodyfat_rpart_qrm_summary <- summary(bodyfat_rpart_qrm) 


#Predict Pruned regression tree model on the bodyfat data set
RegressionTree <- predict(bodyfat_prune, data=bodyfat)

#Create observed value and the predicted value
observed <- bodyfat$DEXfat
predict <- RegressionTree



#Regression.Tree MSE
RegressionTree.MSE <- mean((observed - predict)^2)

#Median Regression MSE
MedianRegression.MSE <- mean(bodyfat_rpart_qrm_summary$residuals^2)

df <- data.frame(
  RegressionTree.MSE,
  MedianRegression.MSE
)

cat("MSE of both models")
df

cat("Based on this, pruned regression tree has lower MSE than median regression model")


cat("Plot based on the regression tree prunning")
plot(as.party(bodyfat_prune), tp_args = list(id=FALSE), main = "Plot based on the regression tree prunning")

cat(
  "Based on the above pruned tree, the variables waist circumference and hip circumference splits explain the majority of the data and I will be choosing one of these variables for quantile regression."
)

cat("Additionally, I will check with linear regression for the variable with significant effect")

check.lm <- lm(DEXfat ~ age + waistcirc + hipcirc + elbowbreadth + kneebreadth, data = bodyfat)

summary(check.lm)

cat("Looks like waistcirc is the most significant, so we will choose this variable")

cat("Now we can run a median quantile regression")
bodyfat_qrm_25 <- rq(DEXfat ~ age + waistcirc, data = bodyfat, tau = 0.25)
summary(bodyfat_qrm_25)
bodyfat_qrm_50 <- rq(DEXfat ~ age + waistcirc, data = bodyfat, tau = 0.50)
summary(bodyfat_qrm_50)
bodyfat_qrm_75 <- rq(DEXfat ~ age + waistcirc, data = bodyfat, tau = 0.75)
summary(bodyfat_qrm_50)

cat("DEXfat explained by waist circumference at quantile 25%, 50%, and 75% regression lines")

plot(data=bodyfat, DEXfat~waistcirc, main="baseR: Quantile regression- DEXfat Explained by waist circumference", xlab='Waist circumference')
abline(rq(DEXfat ~ waistcirc, data=bodyfat, tau = 0.25), col='blue')
abline(rq(DEXfat ~ waistcirc, data=bodyfat, tau = 0.50), col='green')
abline(rq(DEXfat ~ waistcirc, data=bodyfat, tau = 0.75), col='black')
legend('topleft', legend = c('25% Quantile', '50% Quantile', '75% Quantile'),
       fill=c('blue','green','black'))

ggplot(data=bodyfat, aes(x=waistcirc, y=DEXfat)) + geom_point() + stat_quantile(quantiles=c(0.25), method='rq', aes(colour='25%')) + stat_quantile(quantiles=c(0.50), method='rq', aes(colour='50%')) + 
  stat_quantile(quantiles=c(0.75), method='rq', aes(colour='75%')) +
  labs(title="ggplot: Quantile regression- DEXfat Explained by waist circumference", x='Waist circumference', y='DEXfat') +
scale_color_manual(name="Quantile Percent", values = c('25%' = "blue", '50%' = "green", '75%' = "black"))

```

Discussion:

Based on the above pruned tree, the variables waist circumference and hip circumference splits explain the majority of the data and I chose waist circumference for quantile regression. Based on this analysis, pruned regression tree has lower MSE than median regression model. The relationship of Dexfat to Age by Waist Circumference, all three quantiles  regression lines have a positive, and seemingly similar slopes.
    
3. Consider {\textbf{db}} data from the lecture notes (package {\textbf{gamlss.data}}). Refit the additive quantile regression models presented ({\textbf{rqssmod}}) with varying values of $\lambda$ (lambda) in {\textbf{qss}}. How do the estimated quantile curves change?

```{r}
#Q3 - db data from gamlss.data package
data(db)
head(db)
dim(db)
tau <- c(.03, .15, .5, .85, .97)


cat("Parameters: lambda = 0")
rqssmod <- vector(mode = "list", length = length(tau))
db$lage <- with(db, age^(1/3))
for (i in 1:length(tau))
  rqssmod[[i]] <- rqss(head ~ qss(lage, lambda = 0), data = db, tau = tau[i])

gage <- seq(from = min(db$age), to = max(db$age), length = 50)
p <- sapply(1:length(tau), function(i) { predict(rqssmod[[i]], newdata = data.frame(lage = gage^(1/3)))
})

pfun <- function(x, y, ...) {
  panel.xyplot(x = x, y = y, ...)
  apply(p, 2, function(x) panel.lines(gage, x))
  panel.text(rep(max(db$age), length(tau)),
             p[nrow(p),], label = tau, cex = 0.9)
}

xyplot(head ~ age, data = db, main = "Head circumference curve with lambda = 0",
       xlab = "Age (years)", ylab = "Head circumference (cm)", pch = 19,
       scales = list(x = list(relation = "free")),
       layout = c(1, 1), col = rgb(.1, .1, .1, .1),
       panel = pfun)

cat("Parameters: lambda = 1; and tau same as before")

rqssmod <- vector(mode = "list", length = length(tau))
db$lage <- with(db, age^(1/3))
for (i in 1:length(tau))
  rqssmod[[i]] <- rqss(head ~ qss(lage, lambda = 1), data = db, tau = tau[i])

gage <- seq(from = min(db$age), to = max(db$age), length = 50)
p <- sapply(1:length(tau), function(i) { predict(rqssmod[[i]], newdata = data.frame(lage = gage^(1/3)))
})

pfun <- function(x, y, ...) {
  panel.xyplot(x = x, y = y, ...)
  apply(p, 2, function(x) panel.lines(gage, x))
  panel.text(rep(max(db$age), length(tau)),
             p[nrow(p),], label = tau, cex = 0.9)
}

xyplot(head ~ age, data = db, main = "Head circumference curve with lambda=1",
       xlab = "Age (years)", ylab = "Head circumference (cm)", pch = 19,
       scales = list(x = list(relation = "free")),
       layout = c(1, 1), col = rgb(.1, .1, .1, .1),
       panel = pfun)

cat("Parameters: lambda = 20; and tau same as before")

rqssmod <- vector(mode = "list", length = length(tau))
db$lage <- with(db, age^(1/3))
for (i in 1:length(tau))
  rqssmod[[i]] <- rqss(head ~ qss(lage, lambda = 20), data = db, tau = tau[i])

gage <- seq(from = min(db$age), to = max(db$age), length = 50)
p <- sapply(1:length(tau), function(i) { predict(rqssmod[[i]], newdata = data.frame(lage = gage^(1/3)))
})

pfun <- function(x, y, ...) {
  panel.xyplot(x = x, y = y, ...)
  apply(p, 2, function(x) panel.lines(gage, x))
  panel.text(rep(max(db$age), length(tau)),
             p[nrow(p),], label = tau, cex = 0.9)
}

xyplot(head ~ age, data = db, main = "Head circumference curve with lambda=20",
       xlab = "Age (years)", ylab = "Head circumference (cm)", pch = 19,
       scales = list(x = list(relation = "free")),
       layout = c(1, 1), col = rgb(.1, .1, .1, .1),
       panel = pfun)
```

Discussion: 

Here, lambda acts as a shinkage factor which causes the quantiles to become smoother (at higher lambda) rather than becoming wavy or rough with lower values of labda (lambda = 0). So, I found that by increasing the penalty term lambda, which is assigned to the slope of the coefficients, the overall fit of the additive quantile regression model can be smoothen.
 
  
4. Read the paper by Koenker and Hallock (2001), posted on D2L. Write a one page summary of the paper. This should include but not be limited to introduction, motivation, case study considered and findings. 

Answer:

Introduction

In Koenker and Hallock (2001), the authors discuss the utility of quantile regression. In this study, they made a comparison on linear regression and quantile regression analyses.The quantile regression allows large sample groups or population size to into fractals distribution or smaller quantiles represented by a parameter 	\( \tau \) (tau) and maintaining the same 	\( \tau \) for observations above and below the quantile and  minimizing the sum of weighted absolute residuals. They show the utility of quantile regression by looking at various types of data sets such as household food spending on household income and child birthweight.

Motivation

The reason to look at these data sets was to illustrate and explain that in some situations, quantile regression is more useful than linear regression. Quantile regression, which measures the median rather than the mean, will yield more suitable fits to results when a distribution includes outliers that could  force the mean in a specific direction. This was addressed in the paper when referring to the' Quantile Engel Curves' where two different points of high income and low food consumption influenced the least square estimate. Quantile regression could eliminate this type of bias by minimizing the sum of absolute residuals. 


Case study

The Infant Birth weight case study aimed to analyze the association between low birth weights of children and several variables including public policy measures. Infants weighing less than 5 pounds, 9 ounces at birth (2500 grams) were defined as low birth weight. Quantile regression was relevant in this analysis because due to the lower tail of the birth weight distribution, all the least square estimates had been skewed.


Findings

The most notable finding was that, based on the least square calculation, boys were usually born heavier than girls by 100 grams. The figure, however, changes with the quantile of 0.05, where boys were only 45 grams larger and even greater at the quantile of 095, where boys were 130 grams larger. The least squares (linear regression) in this case does a poor job of explaining the distribution variability. Additionally, infants born to black and white mothers with 5th percentile birth weights differed by 1/3 of a kilogram. Furthermore, the age of mother, education beyond high school, prenatal care, marital status, and smoking all contribute to the birth weight of an infant.


Conclusion

In conclusion, some distributions with longer tails can weigh the mean so that the regression of the least squares can provide a false representation of the results. A more precise description of the data and underlying patterns can be obtained using quantile regression. This is particularly useful in econometrics where there are often large outliers that can have a significant impact on a least squares model. Basically, we need to assess our needs when dealing with the datasets: 
    i. Are we concerned about percentiles or median value? 
    ii. Whether we are interested in average values and minimizing the residuals errors. 
